A non-conducting role of the Cav1.4 Ca2+ channel drives homeostatic plasticity at the cone photoreceptor synapse

Version 1

Abstract

Abstract

In congenital stationary night blindness type 2 (CSNB2)—a disorder involving dysfunction of the Cav1.4 Ca2+ channel—visual impairment is relatively mild considering that Cav1.4 mediates synaptic transmission by rod and cone photoreceptors. Here, we addressed this conundrum using a Cav1.4 knockout (KO) mouse and a knock-in (KI) mouse expressing a non-conducting Cav1.4 mutant. Surprisingly, aberrant Cav3 currents were detected in cones of the KI and KO but not wild-type mice. Cone synapses, which fail to develop in KO mice, are present but enlarged in KI mice. Moreover, light responses in cone pathways and photopic visual behavior are preserved in KI but not in KO mice. In CSNB2, we propose that Cav3 channels maintain cone synaptic output provided that the Ca2+-independent role of Cav1.4 in cone synaptogenesis remains intact. Our findings reveal an unexpected form of homeostatic plasticity that relies on a non-canonical role of an ion channel.

Introduction

At the first synapse in the visual pathway, the light-dependent graded electrical signals produced in rod and cone photoreceptors gates the release of glutamate onto postsynaptic neurons. To accomplish this task, photoreceptor synapses are specialized with a ribbon organelle, which helps prime synaptic vesicles 1, 2 and postsynaptic dendrites from horizontal and bipolar cells that invaginate deep within the terminal 3. A variety of proteins interact with the ribbon and synaptic vesicles near release sites (i.e., active zones) 4. The importance of these proteins for vision is illustrated by the numerous inherited retinal diseases linked to mutations in their encoding genes 5.

One such gene is CACNA1F, which encodes the voltage-gated Ca2+ (Cav) channel expressed in retinal photoreceptors, Cav1.4 6–8. Among the sub-family of Cav1.x L-type channels, Cav1.4 exhibits unusually slow inactivation that is well-matched for supporting the tonic, Ca2+-dependent release of glutamate from photoreceptor synaptic terminals in darkness 9,10. More than 200 mutations in CACNA1F cause vision disorders including congenital stationary night blindness type 2 (CSNB2) 11, 12. These mutations are broadly categorized as producing a gain of function or loss of function in Cav1.4 13. How these mutations in CACNA1F lead to the variable clinical phenotypes of CSNB2 is largely unknown. Symptoms may include strabismus, low visual acuity, and in many cases, night blindness 14, 15. The latter suggests a primary defect in rod pathways, which is surprising given that knockout (KO) mice are completely blind and lack any evidence of either rod or cone synaptic responses 6, 7, 16. A major caveat is that rod and cone synapses fail to form in Cav1.4 KO mice 7, 16–18. Thus, Cav1.4 KO mice are not suitable for studies of how CACNA1F mutations differentially affect rod and cone pathways or for efforts to uncover how the biophysical properties of Cav1.4 shape photoreceptor synaptic release properties.

Here, we overcome this hurdle with a knock-in mouse strain (G369i KI) expressing a non-conducting mutant form of Cav1.4 19. We show that cone ribbon synapses in G369i KI mice are largely preserved and that downstream signaling through cone pathways, although greatly impaired, can support visual function. This novel mechanism requires the ability of the Cav1.4 protein, independent of its Ca2+ conductance, to nucleate the assembly of cone ribbon synapses and involves an aberrant Cav3 (T-type) conductance that appears when Cav1.4 Ca2+ signals are compromised.

Results

Ca2+ currents in cones are mediated by Cav3 channels upon Cav1.4 loss-of-function

A prevailing yet unsupported hypothesis regarding the relatively mild visual phenotypes in CSNB2 is that additional Cav subtypes may compensate for Cav1.4 loss of function in cones. If so, then Ca2+ currents (ICa) mediated by these subtypes should be evident in cones of Cav1.4 KO and G369i KI mice. The G369i mutation is an insertion of a glycine residue in a transmembrane domain, which prevents Ca2+ permeation through the channel 19. Rods of G369i KI mice lack any evidence of ICa, despite the normal presynaptic clustering of the mutant channel 19. To test if this might differ in G369i KI cones, we first analyzed the localization of the mutant G369i Cav1.4 channels in cones by immunofluorescence with antibodies against Cav1.4, as well as cone arrestin (CAR) and CtBP2 to label cone terminals (i.e., pedicles) and ribbons, respectively (Fig. 1). In Cav1.4 KO mouse retinas, cone pedicles were shrunken and retracted into the outer nuclear layer (ONL, Fig. 1a) and had malformed ribbons (Fig. 1b). In contrast, cone pedicles in G369i KI mice were normally localized in the outer plexiform layer (OPL, Fig. 1a) and were populated by multiple ribbons (Fig. 1b). As in WT mice, labeling for Cav1.4 was clustered near elongated ribbons in cones of G369i KI mice (Fig. 1b). Thus, unlike in Cav1.4 KO mice, the mutant Cav1.4 protein is normally localized and supports the integrity of cone pedicles and ribbons in G369i KI mice.

Cone pedicles and the mutant Cav1.4 channel are normally localized in the OPL of G369i KI mice but not Cav1.4 KO mice.

Confocal images of the ONL and OPL of WT, G369i KI, and Cav1.4 KO mice labeled with antibodies against cone arrestin (CAR), CtBP2, and Cav1.4. a, Inverted images of CAR labeling. Lower panels correspond to boxed region of the upper panels and depict pedicles labeled by CAR antibodies (dotted outlines). Cone pedicles remain within the OPL of WT and G369i KI retina (arrows) but are misshapen and retracted in the ONL of the CaV1.4 KO retina (arrowheads). b, High magnification, deconvolved images show CaV1.4 labeling near cone ribbons in WT and G369i KI pedicles (arrows) and ribbon spheres without CaV1.4 labeling in the CaV1.4 KO pedicle (arrowheads).

We next compared patch clamp recordings of cones in retinal slices of adult WT, G369i KI, and Cav1.4 KO mice under conditions designed to isolate ICa (Fig. 2). In WT cones, there was a large, sustained ICa that activated around -50 mV and peaked near -20 mV, consistent with the properties of Cav1.4 (Fig. 2a-d). A small-amplitude ICawas detected in cones of G369i KI and Cav1.4 KO mice but activated at significantly more negative voltages than in WT cones (Fig. 2a-d; Table 1). Moreover, ICa in the G369i KI and Cav1.4 KO cones activated around -60 mV and peaked near -35 mV. This aberrant ICa was not sustained as in WT cones but inactivated rapidly during 50-ms (Fig. 2b) and 500-ms (Fig. 2e) step depolarizations. Overlay of the conductance-voltage (Fig. 2d) and inactivation curves (Fig. 2f) revealed a sizeable window current. These features of ICa in G369i KI and Cav1.4 KO cones resembled those of Cav3 T-type channels rather than Cav1.4 20.

A low-voltage activated ICa is present in cones of G369i KI and CaV1.4 KO but absent in WT mice.

a,b, Representative traces of ICaevoked by voltage ramps (a) and voltage steps (b). c,d, I-V (c) and G-V (d) were plotted against test voltage for ICa evoked by 50-ms voltage steps from a holding voltage of -90 mV. Numbers of cells: WT, n = 13; G369i KI, n = 9; CaV1.4 KO, n = 3. e, Representative ICa traces (top) and voltage-protocol (bottom) for steady-state inactivation. ICa was evoked by a conditioning pre-pulse from -90 mV to various voltages for 500 ms followed by a test pulse to -30 mV for 50 ms. f, I/Imax represents the current amplitude of the test pulse normalized to current amplitude of the pre-pulse and was plotted against pre-pulse voltage. Numbers of cells: WT, n = 8; G369i KI, n = 8; Cav1.4 KO, n = 3. In graphs in c,d, and f, smooth lines represent Boltzmann fits, symbols and bars represent mean ± SEM, respectively. In graph in f, line without symbols represents G-V curve for G369i KI cones replotted from d. Shaded region indicates window current for G369i KI cones.

Although Cav3 channels were reported in patch clamp recordings of cone pedicles in WT mouse cones 21, we did not observe a low voltage-activated component of ICa in the I-V curve from our recordings of WT mouse cone somas, which would be indicative of a Cav3 subtype (Fig. 2a,c). Moreover, ICa in WT cones was blunted by the Cav1 antagonist isradipine (Fig. 3a) but not the Cav3 antagonist, ML 218 (Fig. 3b; ML 218 caused a minor suppression of ICa in some WT cones, which could be attributed to weak activity on Cav1.4, Supp.Fig. S1). In contrast, ICain G369i KI cones was significantly suppressed by ML 218 but not by isradipine (Fig. 3a,b). To further test for a Cav3 contribution to ICa, we analyzed cones of ground squirrel retina, where the large amplitude of ICa facilitates pharmacological analyses. Consistent with its actions on Cav1.4 in WT mouse cones (Fig. 3c) and transfected HEK293T cells (Supp.Fig. S1), ML 218 caused an insignificant inhibition of peak ICa (-7.0 ± 20.8%, n = 6 cones) as well as a negative shift in the voltage-dependence of activation in ground squirrel cones (Supp.Fig. S2). As a positive control, we confirmed that ML 218 blocked a prominent Cav3-type current in ground squirrel type 3a OFF cone bipolar cells (Supp.Fig. S2). Application of isradipine followed by ML 218 resulted in a time-and voltage-dependent suppression of the residual ICa, which is a hallmark of Cav1 inhibition by dihyropyridine antagonists such as isradipine (Fig. 3d, Supp.Fig. S2)22. Finally, we also performed patch clamp recordings of cone terminals in macaque retina where there was no evidence of a Cav3 current (Supp.Fig. S3). We conclude that Cav3 channels contribute significantly to ICa in cones only when Cav1.4 Ca2+ signals are absent.

Pharmacological characterization of ICa in cones of WT and G369i KI mice and ground squirrel.

a,b, Analysis of mouse cones. Left, representative traces for ICa evoked by voltage ramps in cones of WT or G369i KI mice before (baseline) and after exposure to 1 μM of isradipine (ISR, a, WT, n = 4; G369i KI, n = 5) or 5 μM ML 218 (b, WT, n = 5; G369i KI, n = 5). Right, ICa amplitudes before (-) and during (+) perfusion of the blocker on the same cells. Each point represents a different cell. *, p < 0.05 by paired t-test. c,d, Analysis of ground squirrel cones. Representative traces corresponding to baseline-corrected ICa evoked by voltage ramps (c,d, left) and corresponding G-V plot (c, right) before and during application of ML 218 (c) or ISR alone or ISR+ML218 (d). In d, ICa amplitudes are plotted before (-) and after (+) block for cones in the superior (sup.; n = 7 cones), middle (mid.; n = 5 cones), and inferior (inf.; n = 5 cones) thirds of the retina. The ICa blocked by ISR alone was measured at the peak of the control I-V curve between -40 and -20 mV. The ICa amplitude blocked by adding ML218 was measured as the average current between -50 and -45 mV before ML218 addition relative to zero current following the addition of ML218. The changes produced by ML218 were small but nonetheless significant (sup., n = 7, ISR: p < 0.0001, ML218: p = 0.0047; mid., n = 5, ISR: p < 0.0001, ML218: p = 0.0001; inf., n =5, ISR: p = 0.0019, ML218: p = 0.0460; two-tailed t test).

Cone synaptogenesis relies on the Cav1.4 protein but not its Ca2+ conductance

As shown in previous studies 7, 16–18, CtBP2 labeled stubby, sphere-like structures resembling immature ribbon material in Cav1.4 KO mice (Fig. 1). The presence of elongated ribbons in G369i KI mice suggested that, as in rods 19, the non-conducting mutant Cav1.4 protein may support the molecular assembly of cone synapses. To this end, we analyzed cone synapses by immunofluorescence and confocal microscopy. Along with the major constituents of the ribbon, CtBP2 and RIBEYE 23, presynaptic proteins such as bassoon and members of the postsynaptic signaling complex in depolarizing (ON) cone bipolar cells (GPR179, mGluR6, and TRPM1 24) were enriched near cone ribbons in G369i KI mice as in WT mice (Fig. 4a). Compared to WT mice, the labeled structures occupied a larger volume (Fig. 4b,c), albeit at a much lower density (Fig. 4d), which increased linearly with the volume of the pedicle in G369i KI mice (Fig. 4e-i), perhaps as a homeostatic response to Cav1.4 loss-of-function.

Immunofluorescence characterization of cone synapses in WT and G369i KI mice.

a, Confocal images of the OPL of WT and G369i KI mice labeled with antibodies against cone arrestin (CAR) and proteins that are presynaptic (CtBP2, bassoon) or postsynaptic (GPR179, TRPM1, mGluR6). Every other panel shows high-magnification, deconvolved images of single pedicles labeled with cone arrestin (rod spherule-associated signals were removed for clarity). Arrows indicate ribbon synapses, which appear enlarged in the G369i KI pedicles. b-d, Violin plots represent volume occupancy of labeling for each synaptic protein normalized to their respective CAR-labeled pedicles. e-i, Dependence of synapse size on pedicle size. Volumes corresponding to labeling of CtBP2 (e: p = 0.051, r = 0.4 for WT; p < 0.0001, r = 0.88 for G369i KI), CaV1.4 (f: p = 0.8, r = 0.06 for WT; p = 0.002, r = 0.88 for G369i KI), bassoon (g: p = 0.1, r = 0.34 for WT; p < 0.0001, r = 0.75 for G369i KI), mGluR6 (h: p = 0.32, r = 0.35 for WT; p = 0.002, r = 0.73 for G369i KI) and TRPM1 (i: p = 0.32, r = 0.35 for WT; p = 0.007, r = 0.66 for G369i KI) are plotted against pedicle volume. Dashed and solid lines represent fits by linear regression for WT and G369i KI, respectively.

The cone synapse is structurally complex, with the dendritic tips of two horizontal cells and an intervening ON cone bipolar cell invaginating deeply into the pedicle near the ribbon 3. To test how the switch in Cav subtypes might affect this arrangement of postsynaptic partners, we generated 3D reconstructions of cone synapses by serial block-face scanning electron microscopy (SBFSEM; Fig. 5). As with the enlarged synaptic contacts (Fig. 4), there was some evidence of structural modifications in G369i KI pedicles. Compared to WT pedicles, ribbons appeared disorganized in G369i KI pedicles which extended telodendria laterally rather than apically (Fig. 5a). In addition, a slightly larger fraction of synaptic sites in the G369i KI pedicle (35% vs 14% in WT) formed incorrect postsynaptic partnerships (Fig.5b-d’; Table 2). Nevertheless, the number of ribbons is normal in G369i KI cones and about half of the ribbons making invaginating contacts with the appropriate cell types (i.e., both horizontal cells and cone bipolar cells; Table 2). Therefore, while necessary for the structural refinement of the cone synapses, Cav1.4 Ca2+ signals are largely dispensable for cone synapse assembly.

Cone synapses form incorrect pairings with postsynaptic partners in G369i KI mice.

3D reconstructions of WT and G369i KI pedicles (n = 2 each) were obtained by SBFSEM. a, 3D renderings showing ribbons (magenta) within cone pedicles (gray) from WT and G369i KI mice. b,b’, 3D renderings (b) show ribbon sites in a WT cone pedicle contacting one horizontal (HC1; yellow) and three bipolar cells (BC1-3; purple). The raw image (b’) shows a single plane example of BC1-2 and HC1 contacting the ribbon site. c-d’, Single plane raw images (left panels, c; d) and 3D reconstructions (right panels, c; d’) show ribbon sites within the G369i KI cone pedicle contacting in c: horizontal cells (HC1-2) only, CBCs only (BC1-2); and in d,d’: glia (the G369i KI cone pedicle contacts a glial cell (orange) and an unknown partner); the glial cell completely envelops the pedicle. Inset in d’ shows glial-contacting ribbon site (arrow). In other panels, arrows indicate points of contact between ribbons and other postsynaptic elements.

Cone signaling to a postsynaptic partner is intact in G369i KI mice

The preservation of cone synapses in G369i KI mice allowed the unique opportunity to test whether a Cav subtype other than Cav1.4 could support ribbon-mediated synaptic release. To this end, we compared synaptic transmission between cones and horizontal cells (HCs, Fig. 6) in retinal slices of WT and G369i KI mice. In darkness, glutamate released from cones depolarizes horizontal cells via activation of α-amino-3-hydroxy-5-methyl-4-isoxazolepropionic (AMPA)/kainate receptors 25, 26. The resulting excitatory postsynaptic current (EPSC) undergoes a decline in response to light stimuli that hyperpolarize cones 27. In WT HCs, a 1 s light pulse (λ=410 nm) inhibited the standing EPSC, which is reflected as an outward (hyperpolarizing “ON”) current (Fig. 6a). Upon termination of the light pulse, an inward (depolarizing “OFF”) current signaled the resumption of the EPSC (Fig. 6a). Both the ON and OFF components of the EPSCs in WT HCs increased with light intensity, reflecting the impact of luminance on the presynaptic membrane potential of cones and the subsequent change in glutamate release from their terminals (Fig. 6b). The ON and OFF components of EPSCs in G369i KI HCs were measurable, although lower in amplitude than in WT (Fig. 6a,b). The EPSC in darkness and the light response were abolished by the AMPA/kainate receptor antagonist, DNQX, in both WT and G369i KI horizontal cells (Fig. 6c,d). Thus, while greatly impaired, cone synapses can transmit light information to postsynaptic partners in the retina of G369i KI mice.

Light responses of horizontal cells are impaired in G369i KI mice.

a-b, Representative traces from whole-cell patch clamp recordings of horizontal cells held at -70 mV (a) and quantified data (b) for currents evoked by 1 s pulses of light (λ = 410 nm) plotted against light intensities. In b, peak current amplitudes during (ON current) and after (OFF current) the light stimuli were plotted against photon flux per μm2 (Φq/μm2). Data represent mean ± SEM. WT, n = 8; G369i KI, n =9. c-d, Representative traces from horizontal cells held at -70 mV (c) and quantified data (d) for currents evoked by 1-s pulses of light (λ = 410 nm, 1.2 x 105 Φq/μm2) before, during, and after washout of DNQX (20 μM). In d, symbols represent responses from individual cells, n = 5 cells for WT and 3 cells for G369i KI, bars represent mean ± SEM. **, p < 0.01 by paired t-tests.

Light responses of bipolar cells and visual behavior is spared in G369i KI but not Cav1.4 KO mice

While horizontal cells act as inhibitory interneurons that modulate photoreceptor output, the vertical dissemination of visual information from cones to the inner retina is relayed by glutamate to ON and OFF cone bipolar cells (CBCs) which express mGluR6 and AMPA/kainate receptors, respectively. To test whether cone synaptic transmission to CBCs is intact in G369i KI mice, we recorded electroretinograms (ERGs) under light-adapted conditions using Cav1.4 KO mice as a negative control (Fig. 7a-d). In these recordings, the light-induced response of photoreceptors and the postsynaptic response of ON CBCs corresponds to the a-and b-waves, respectively. Unlike in Cav1.4 KO mice, the a-waves of G369i KI mice were like those in WT mice, which indicates that cones do not degenerate in this mouse strain. While reduced in amplitude, the b-wave was measurable in G369i KI mice and significantly larger than in Cav1.4 KO mice at the highest light intensities (Fig. 7a,b). We also recorded flicker ERGs using 10 Hz light stimuli that can isolate cone pathways involving both ON and OFF CBCs. In WT mice, flicker responses exhibited two peaks, one at a lower irradiance (-2 log cd·s/m2) and one at higher irradiance (0.5 log cd·s/m2). The peak at the lower irradiance is attributed to responses in both rod and cone pathways, and the peak at higher irradiance is attributed to responses exclusively in cone pathways (Fig. 7c,d)28, 29. While significantly lower in amplitude than in WT mice, flicker responses in G369i KI mice showed a similar non-monotonic relation. Compared to Cav1.4 KO mice, G369i KI mice showed peak flicker responses that did not differ at -2 log cd·s/m2, but were significantly higher at 0.5 log cd·s/m2 (Fig.7d). The inverted flicker responses at higher illuminations in G369i KI mice (Fig. 7c) were absent in Cav1.4 KO mice and may result from the hyperpolarizing contribution of cone-to-OFF CBC transmission. These results suggest that cone-to-CBC signaling is intact in G369i KI mice.

Photopic vision is reduced but not lost in G369i KI mice.

a, Representative traces of photopic ERGs recorded in the presence of background green light (20 cd · s/m2) in WT, G369i KI and CaV1.4 KO mice. Flash intensities are shown at left. Arrows and arrowheads depict the a-and b-waves, respectively. b, a-wave (left) and b-wave (right) amplitudes are plotted against light intensity. Symbols and bars represent mean ± SEM. WT, n = 7; G369i KI, n = 5; CaV1.4 KO, n = 6. For a-waves, there was a significant effect of light intensity (p < 0.001) and genotype (p = 0.0086) by a mixed-effects model; by post-hoc Tukey test, there was no significant difference between WT and G369i KI at any light intensity. For b-waves at the highest light intensity, WT vs G369i KI, p = 0.0001; WT vs. CaV1.4 KO, p < 0.0001; G369i KI vs. CaV1.4 KO, p = 0.0038, Two-way ANOVA with Tukey’s post hoc analysis. c,d, Representative traces (c) and quantified data (d) for 10 Hz flicker responses evoked by white light flashes of increasing luminance (from -4 to 2 log cd· s/m2). Arrows in c depict inverted waveform responses in G369i KI mice that are absent in CaV1.4 KO mice. Symbols and bars represent mean ± SEM. WT, n = 4; G369i KI, n = 5; CaV1.4 KO, n = 5. At each of the 3 highest light intensities there was a significant difference (p < 0.05) in b-waves of WT vs G369i KI and G369i KI vs Cav1.4 KO by Two-way ANOVA with Tukey’s post hoc analysis. e, Representative swim path traces of WT, G369i KI and CaV1.4 KO mice from the visible platform swim tests performed in the dark (upper traces) and light (lower traces). f, Quantified latency to platform. Symbols represent the average of the last 3 swim trials for each mouse of each genotype for both dark and light conditions. Dotted lines represent the mean. WT, n = 11; G369i KI, n = 10; CaV1.4 KO, n = 9. *, p < 0.05; ***, p < 0.001; ****, p < 0.0001; Kruskal-Wallis one-way ANOVA with Dunn’s post hoc analysis.

To validate these results with respect to vision-guided behavior, we used a swim test that assesses the ability of mice to identify a visible platform 30. In darkness, G369i KI and Cav1.4 KO mice took significantly longer to find the platform compared to WT mice (Fig. 7e,f). These results are consistent with our flicker response assays (Fig.7c,d) as well as the absence of ICa in rods and rod-to-rod bipolar cell synaptic transmission in G369i KI mice 19. In daylight conditions, G369i KI mice but not Cav1.4 KO mice performed as well as WT mice (Fig. 7e,f). Thus, G369i KI mice are unique in retaining visual function under photopic but not scotopic conditions. Collectively, our results suggest that Cav3 channels can support cone synaptic responses and visual behavior in G369i KI but not Cav1.4 KO mice.

Discussion

The nervous system has remarkable abilities to adapt to pathological perturbations in neuronal activity. In the retina, ablation or degeneration of photoreceptors triggers various postsynaptic mechanisms that maintain some level of visual function in rod or cone pathways. These include remodeling of bipolar cell dendrites and their synapses31–33 as well as changes in the sensitivity of bipolar cells to photoreceptor input and inhibitory modulation33, 34. To our knowledge, this study provides the first evidence for a presynaptic form of homeostatic plasticity that originates within photoreceptors. Using Cav1.4 KO and G369i KI mice, we identify the upregulation of a Cav3 conductance as a common response to Cav1.4 loss-of-function in cones. However, Cav3 channels can only compensate for Cav1.4 loss-of-function when cone synapse structure is maintained. Thus, our results also highlight a crucial, non-conducting role for the Cav1.4 protein that allows cone synapses to function in the absence of Cav1.4 Ca2+ signals.

A non-canonical role for Cav1.4 in regulating cone synapse assembly

A major finding of our study is that cone synapse formation requires the Cav1.4 protein but not Cav-mediated Ca2+ influx. As shown for rod synapses 19, ribbons and other components of the pre-and post-synaptic complex assemble normally at cone synapses in G369i mice (Figs. 1,4,5). Cav3 Ca2+ signals are dispensable for this process since their presence in Cav1.4 KO cones is not accompanied by any semblance of ribbon synapses (Fig.1B). An intimate relationship between Cav1.4 and ribbons is supported by the colocalization of Cav1.4 and RIBEYE puncta resembling ribbon precursor spheres in the developing OPL7. Moreover, light adaptation, which decreases the size of the ribbon, leads to a reduction in Cav1.4 labeling in mouse retina2. While evidence for the binding of RIBEYE to Cav1.4 is lacking, such an interaction could support the oligomerization of RIBEYE A and B domain within a macromolecular complex 35. Alternatively, Cav1.4 could pioneer sites of ribbon assembly perhaps by serving as a docking or nucleation site for the active zone. Regardless of the mechanism, our findings show that the formation of photoreceptor synaptic complexes does not require Ca2+ influx through Cav1.4 channels.

Despite a normal molecular organization, G369i KI cone synapses were enlarged and made errors in postsynaptic partner selection. While I-V curves predict similar peak ICa amplitudes in cones of WT and G369i KI mice near the membrane potential of cones in darkness (-45 to -50 mV 36, Fig.1c), the strong inactivation of Cav3 channels predicts greatly diminished Ca2+ signals in G369i KI pedicles. Paradoxically, reductions in presynaptic Ca2+ in rod photoreceptors are thought to cause illumination-dependent shrinkage of ribbons in vivo and in vitro 2, 37, 38. However, an inverse correlation between presynaptic Ca2+ influx and ribbon synapse size is seen in sensory hair cells in zebrafish 39. The mechanism involves decreased mitochondrial Ca2+ uptake and increase in the redox state of nicotinamide adenine dinucleotide (NAD+/NADH ratio) 40. If a similar mechanism applied in cones, Cav3 Ca2+ signals in G369i KI cones may decay too quickly to enable mitochondrial Ca2+ uptake mechanisms that trim ribbons. The size of ribbons and the postsynaptic specialization increased linearly with pedicle volume in G369i KI but not in WT mice (Fig. 4e-i), which could represent a form of homeostatic synaptic scaling as has been demonstrated at the neuromuscular junction 41, 42. Similarly, the presence of some non-invaginating contacts with incorrect partner pairings in G369i KI mice (Fig.5, Table 2) could represent a compensatory response to weakened synaptic output. Considering that only a subset of CBC subtypes re-wire correctly following partial ablation of cones in immature mice 32, the structurally normal cone synapses of G369i KI mice could involve contacts with CBCs that are unusually resilient to loss of presynaptic input.

While it is well-established that Cav1.4 is the main Cav subtype in mouse cones, Cav3 channels, in particular Cav3.2, have been reported to be expressed in mouse cones by electrophysiology 21 and single cell RNA-seq studies 21, 43, 44. By pharmacological and other criteria, we found no evidence for a functional contribution of Cav3 in our recordings of cones in WT mice (Figs. 2,3), ground squirrels, or macaque (Supp.Figs. S2,S3). A caveat of using dihydropyridine antagonists such as isradipine to isolate the contribution of Cav1 from other Cav subtypes is the relatively low sensitivity of Cav1 channels to these drugs and their strong voltage-dependence. At a holding voltage of -90 mV, isradipine at 1 μM causes only ∼80% inhibition of Cav1.4 with greater block at depolarized voltages 22, which agrees with our recordings of cones in WT mice (Fig.3a) and ground squirrels (Fig.3d, Supp.Fig.S2). Moreover, Cav3 blockers such as Z944 45 and ML218 (Supp.Fig. S2) have additional activity on Cav1 channels at micromolar concentrations including effects on current amplitude and activation voltage. Thus, currents mediated by Cav1.4 that are sensitive to Cav3 blockers (e.g, >5 µM ML218) and spared by dihydropyridines at negative voltages may be mistaken as being mediated by Cav3. In our experiments, the biophysical properties of the isradipine-sensitive ICain cones of WT mice and ground squirrels resembled only those of Cav1, whereas those of the ML218-sensitive ICa in cones of G369i KI mice resembled only those of Cav3. Therefore, we favor the interpretation that Cav3 contributes to ICa in mouse cones only upon silencing of the Cav1.4 Ca2+ conductance.

Consistent with the diminutive Cav3-mediated ICa, light responses of HCs were evident but smaller in G369i KI than in WT mice. Due to their slow activation and strong inactivation 46, Cav3 channels may have a reduced ability to fuel the Ca2+ nanodomains that support fast and sustained components of release, both of which occur only at ribbon sites in cones1 47. Cav3 channels may also be located further from the ribbon than Cav1.4, thus lowering the efficiency of coupling to exocytosis. Unfortunately, it was not possible to test this with commercially available antibodies against Cav3.2, which yielded identical patterns of immunofluorescence in the OPL of WT and Cav3.2 KO mice (data not shown). A detailed analysis of the Cav3 subtype(s) and their subcellular localization in G369i KI cones is required to unravel the shortcomings of Cav3 channels with respect to cone synaptic release.

Based on extremely heterogeneous clinical presentations, CSNB2 manifests as a spectrum of visual disorders that originate from various mutations in CACNA1F 14, 48. Even though some CSNB2 mutant Cav1.4 channels may traffic normally to the plasma membrane, many of these mutations are expected to produce non-functional, non-conducting Cav1.4 channels.49, 50 Yet, the visual phenotypes of CSNB2 patients are not as severe as the complete blindness in Cav1.4 KO mice, which lack any Cav1.4 protein expression and exhibit no signs of visual behavior (Fig.7e,f)51. Collectively, our results suggest that G369i KI mice accurately model Cav1.4 channelopathies in CSNB2 patients that are characterized by a greater impairment in rod than in cone pathways. This interpretation is supported by our findings that G369i KI mice exhibit horizontal cell responses to bright but not dim illumination (Fig. 6), ERG responses under conditions of light adaptation (Fig.7a-d) but not dark-adaptation 19, and visual behavior under photopic but not scotopic conditions (Fig. 7d). Together with the enlargement of synaptic sites, modest levels of synaptic release from cones of G369i KI mice may be sufficient to support nominal transmission of visual information through cone pathways. Furthermore, we acknowledge that G369i KI mice could also exhibit homeostatic alterations in the inner retina, which are known to support visual function when photoreceptor input is severely compromised 32–34. Future studies of the retinal circuitry and visual behavior of G369i KI mice could identify compensatory pathways that are recruited upon Cav1.4 loss-of-function and how they might be targeted in novel therapies for CSNB2 and related disorders.

eLife Assessment

Based on analyses of retinae from genetically modified mice, and from wild-type ground squirrel and macaque employing microscopic imaging, electrophysiology, and pharmacological manipulations, this useful study on the role of Cav1.4 calcium channels in cone photoreceptor cells (i) shows that the expression of a Cav1.4 variant lacking calcium conductivity supports the development of cone synapses beyond what is observed in the complete absence of Cav1.4, and (ii) indicates that the cone pathway can partially operate even without calcium flux through Cav1.4 channels, thus preserving behavioral responses under bright light. The evidence for the function of Cav1.4 protein in synapse development is convincing, and in agreement with a closely related earlier study by the same authors on rod photoreceptors, but the evidential support for the notion of a homeostatic compensation of Cav1.4 loss by Cav3 is incomplete. As congenital Cav1.4 dysfunction can cause stationary night blindness, this work relates to a wide range of neuroscience topics, from synapse biology to neuro-ophthalmology.

Reviewer #1 (Public Review)

Cav1.4 calcium channels control voltage-dependent calcium influx at photoreceptor synapses, and congenital loss of Cav1.4 function causes stationary night blindness CSNB2. Based on a broad portfolio of methodological approaches - genetic mouse models, immunolabeling and microscopic imaging, serial block-face-SEM, ERGs, and electrophysiology - the authors show that cone photoreceptor synapse development is strongly perturbed in the absence of Cav1.4 protein, and that expression of a nonconducting Cav1.4 channel mitigates these perturbations. Further data indicate that Cav3 channels are present, which, according to the authors, may compensate for the loss of Cav1.4 calcium currents and thus maintain cone synaptic transmission. These data, which are in agreement with a similar study by the same authors on rod photoreceptor synapses, help to explain what functional defects exactly cause CSNB2 and why it is accompanied by only mild visual impairment.

The strengths of the present study are its conceptual and experimental soundness, the broad spectrum of cutting-edge methodological approaches pursued, and the convincing differential analysis of mutant phenotypes.

Weaknesses mainly concern the experiments and arguments leading to the authors' notion that Cav3 channels may partially compensate for the loss of Cav1.4 calcium currents in cone synapses. It is possible that the non-conducting Cav1.4 variant supports synapse development and the Cav3 channel then provides the calcium influx. However, in its current state, the study does not unequivocally assess Cav3 expression in wild-type cones, it lacks direct evidence of Cav3 expression and upregulation, e.g. via single cell transcriptomics, immunolabeling, or an elaboration on electrophysiology, and it does not test the authors' earlier idea that Cav1.4 might couple to intracellular calcium stores at photoreceptor synapses

Reviewer #2 (Public Review)

Summary:

This paper by Maddox et al. presents the results of a study of Ca channel function in mouse cone photoreceptor synaptic terminals. It builds on earlier work by the same authors (Maddox et al. 2020 in eLife) which demonstrated that a non-conducting but voltage-sensing variant of Cav1.4 (G369i knock-in, or KI) could substitute for WT Cav1.4 to promote relatively normal rod synapse development despite an inability to support Ca2+-dependent glutamatergic transmission to postsynaptic bipolar cells. Cav1.4 knock-out (KO) rod synapses, however, were completely disorganized, indicating that the presence of Cav1.4 protein is critical for synaptic organization. Here, the authors extend their study of the G369i-KI retina to demonstrate that G369i-KI cones develop working (though disrupted and sometimes aberrant) synapses that support some visual function owing to compensatory expression of Cav3-containing Ca channels that can mediate some Ca2+-dependent transmission from cones to postsynaptic cells. This compensatory expression of a low voltage-activated Ca conductance was not noted previously (Maddox et al. 2020) in G369i-KI rods.

Strengths:

In all, this is a scientifically sound study that shows obvious differences between synaptic terminal morphology and organization, macroscopic Ca currents, transmission to postsynaptic horizontal and bipolar cells (with whole-cell recording and ERG, respectively), and visually-guided behavior in experimental groups.

Weaknesses:

The major criticism that I have of the study is that it infers Ca channel molecular composition based solely on pharmacological analysis, which, as the authors note, is confounded by the cross-reactivity of many of the "specific" channel-type antagonists. The authors note that Cav3 mRNAs have been found in cones, but here, they do not perform any analysis to examine Cav3 transcript expression after G369i-KI nor do they examine Ca channel transcript expression in monkey or squirrel cones, which serve as controls of sorts for the G369i-KI (i.e. like WT mouse cones, cones of these other species do not seem to exhibit LVA Ca currents).

Secondarily, in Maddox et al. 2020, the authors raise the possibility that G369i-KI, by virtue of having a functional voltage-sensing domain-might couple to intracellular Ca2+ stores, and it seems appropriate that this possibility be considered experimentally here.

As a minor point: the authors might wish to note - in comparison to another retinal ribbon synapse-that Zhang et al. 2022 (in J. Neuroscience) performed a study of mouse rod bipolar cells found a number of LVA and HVA Ca conductances in addition to the typical L-type conductance mediated by Cav1-containing channels.

Reviewer #3 (Public Review)

Summary

This is an important study that tests the hypothesis that Cav1.4 calcium channels do more than provide a voltage-dependent influx of Ca2+ into photoreceptors. The relevant background can be divided into two tranches. First, deletion of Cav1.4 channels (Cav1.4 knock-out) disrupts rod and cone photoreceptors and their synapses in the outer plexiform layer. Second, knock-in of a non-conducting Cav1.4 channel (Cav1.4 knock-in) partially spares the organization of the outer plexiform layer and photoreceptor synapses (Maddox et al., eLife 2020), which is remarkable considering the disruption of the outer plexiform layer in the Cav1.4 knock-out. In addition, phototransduction, assessed by scotopic and phototopic electroretinography (a-wave amplitude) in the Cav1.4 knock-in retina was partially spared for rods and only slightly impaired for cones. However, the non-conducting Cav1.4 channel of the Cav1.4 knock-in failed to rescue synaptic transmission across the outer retina (electroretinography: b-wave amplitude, Maddox et al., eLife 2020). The 2020 Maddox et al. (eLife) focused more on the rod pathway, while the current work addressed the cone pathway.

Strengths

The study addresses the important question of how disruption of Cav1.4 function in both rod and cone photoreceptors leads to impairment primarily of the rod pathway for scotopic vision. This is clinically relevant as human mutations lead to stationary night blindness rather than blindness. The work relevance provides excellent single-cell electrophysiological recordings of Ca2+ currents from cones of wild-type, Cav1.4 knock-out, and Cav1.4 knock-in mice and, in addition, from ground squirrel and monkey cones. To make these recordings successfully in the various species and the compromised retinas (Cav1.4 knock-out and Cav1.4 knock-in) is very impressive. The findings clearly advance our understanding of Ca2+ channel function in cones. In addition, the study presents high-quality electron microscopy reconstructions of cones and further physiological and behavioral data related to the cone pathway.

Weaknesses

The major critiques are related to the description of the Cav1.4 knock-in mouse as "sparing" function, which can be remedied in part by a simple rewrite, and in certain places, the data may need to be examined more critically. In particular, the authors should address features in the data presented in Figures 6 and 7 that seem to indicate that the retina of the Cav1.4 knock-in is not intact, but the interpretation given by the authors as "intact" is not appropriate and made without rigorous statistical testing.

Reviewer #4 (Public Review)

Summary:

Cav1.4 voltage-gated calcium channels play an important role in neurotransmission at mammalian photoreceptor synapses. Mutations in the CACNA1f gene lead to congenital stationary night blindness that particularly affects the rod pathway. Mouse Cav1.4 knockout and Cav1.4 knockin models suggest that Cav1.4 is also important for the cone pathway. Deletion of Cav1.4 in the knockout models leads to signaling malfunctions and to abundant morphological re-arrangements of the synapse suggesting that the channel not only has a role in the influx of Ca2+ but also in the morphological organization of the photoreceptor synapse. Of note, also additional Cav-channels have been previously detected in cone synapses by different groups, including L-type Cav1.3 (Wu et al., 2007; pmid; Kersten et al., 2020; pmid), and also T-type Cav3.2 (Davison et al., 2021; pmid 35803735).

In order to study a conductivity-independent role of Cav1.4 in the morphological organization of photoreceptor synapses, the authors generated the knockin (KI) mouse Cav1.4 G369i in a previous study (Maddox et al., eLife 2020; pmid 32940604). The Cav1.4 G369i KI channel no longer works as a Ca2+-conducting channel due to the insertion of a glycine in the pore-forming unit (Madox et al. elife 2020; pmid 32940604). In this previous study (Madox et al. elife 2020; pmid 32940604), the authors analyzed Cav1.4 G369i in rod photoreceptor synapses. In the present study, the authors analyzed cone synapses in this KI mouse.

For this purpose, the authors performed a comprehensive set of experimental methods including immunohistochemistry with antibodies (also with quantitative analyses), electrophysiological measurements of presynaptic Ca2+ currents from cone photoreceptors in the presence/absence of inhibitors of L-type- and T-type- calcium channels, electron microscopy (FIB-SEM), ERG recordings and visual behavior tests of the Cav G369i KI in comparison to the Cav1.4 knockout and wild-type control mice.

The authors found that the non-conducting Cav channel is properly localized in cone synapses and demonstrated that there are no gross morphological alterations (e.g., sprouting of postsynaptic components that are typically observed in the Cav1.4 knockout). These findings demonstrate that cone synaptogenesis relies on the presence Cav1.4 protein but not on its Ca2+ conductivity. This result, obtained at cone synapses in the present study, is similar to the previously reported results observed for rod synapses (Maddox et al., eLife 2020, pmid 32940604). No further mechanistic insights or molecular mechanisms were provided that demonstrated how the presence of the Cav channels could orchestrate the building of the cone synapse.

Strengths:

The study has been expertly performed. A comprehensive set of experimental methods including immunohistochemistry with antibodies (also with quantitative analyses), electrophysiological measurements of presynaptic Ca2+ currents from cone photoreceptors in the presence/absence of inhibitors of L-type- and T-type- calcium channels, electron microscopy (FIB-SEM), ERG recordings and visual behavior tests of the Cav G369i KI in comparison to the Cav1.4 knockout and wild-type control mice.

Weaknesses:

The study has been expertly performed but remains descriptive without deciphering the underlying molecular mechanisms of the observed phenomena, including the proposed homeostatic switch of synaptic calcium channels. Furthermore, a relevant part of the data in the present paper (presence of T-type calcium channels in cone photoreceptors) has already been identified/presented by previous studies of different groups (Macosko et al., 2015; pmid 26000488; Davison et al., 2021; pmid 35803735; Williams et al., 2022; pmid 35650675). The degree of novelty of the present paper thus appears limited.

Author response:

The following is the authors’ response to the original reviews.

Reviewer #1 (Recommendations For The Authors):

The authors should perform experiments to answer this question: does Cav3 transcription increase in the G369i-KI, or is there instead some post-transcriptional modulation that permits surface expression of functional Cav3-containing channels in the absence of typical HVA Ca conductances? Also, the authors should determine whether G369i-KI can mediate Ca2+ release from intracellular stores and whether release from stores is upregulated as Cav3-containing channel expression (or function) is increased.

We performed transcriptomic (drop-seq) analysis to test whether a Cav3 subtype is upregulated in cones of G369i KI mice. These experiments show that, consistent with previous studies (PMID 35803735, 26000488), Cacna1h appears to be the primary Cav3 subtype expressed mouse cones. However, as shown in new Supp.Fig.S3, there was no significant difference in the levels of Cacna1h transcripts in WT and G369i KI cones. Therefore, we propose that there may be some post-transcriptional modification, or alteration in a pathway that regulates channel availability, that enables the contribution Cav3 channels to the whole-cell Ca2+ current in the absence of functional Cav1.4 channels cones.

We also performed Ca2+ imaging experiments in WT vs G369i KI cone terminals to assess whether the diminutive Cav3 current in G369i KI cone terminals may be compensated by upregulation of a Ca2+ signal such as from intracellular stores. Arguing against this possibility, depolarization-evoked Ca2+ signals in G369i KI cones were dramatically reduced compared to WT cones (new Fig.9).

Reviewer #2 (Recommendations For The Authors):

Major points-

(1) It is stated in too many places that cone features in the Cav1.4 knock-in are "intact", preserved, or spared, but this representation is not accurate. There are two instances in this study that qualify as intact when comparing KI to WT: 1) the photopic a-waves in the Cav1.4 knock-in (also demonstrated in Maddox et al 2020) and 2) latency to the platform (current MS, Figure 7f). However, in the numerous instances listed below, the authors compared the Cav1.4 knock-in to the Cav1.4 knock-out, and then referred to the KI as exhibiting intact responses. The reference point for intactness needs to be wildtype, as appropriately done for Figures 2 and 3, and when comparing the KI to the KO the phrasing should be altered; for example: "the KI was spared from the extensive degeneration witnessed in the KO....".

In most cases, we clearly note that there are key differences in the WT and the G369i KI cone synapses, which highlight the importance of Cav1.4-specific Ca2+ signals for certain aspects of the cone synapse. We disagree with the reviewer on the point that we did not often use the WT as a reference since most of our experiments involved comparisons of only WT and G369i KI (Figs. 3-6) or WT, G369i KI, and Cav1.4 KO (Figs.1,7—and in these cases comparisons specifically between WT and G369i KI mice were included). We used “intact” as a descriptor for G369i KI cone synapses since these are actually present, albeit abnormal in the G369i KI retina, whereas cone synapses are completely absent in the Cav1.4 KO retina. To avoid confusion, we modified our use of “intact” and “preserved” where appropriate.

A. Abstract, line 34 to 35: ".......preserved in KI but not in KO.".

Abstract was rewritten and this line was removed.

B. Line 36: "....synaptogenesis remains intact". The MS documents many differences in the morphology of KI and WT cones (immunofluorescence and electron microscopy data), which is counter to an intact phenotype.

The sentence was: “In CSNB2, we propose that Cav3 channels maintain cone synaptic output provided that the Ca2+-independent role of Cav1.4 in cone synaptogenesis remains intact.”

Here the meaning of “intact” refers to the Ca2+ -independent role of Cav1.4, not synapses. Thus, we have left the sentence unchanged.

C. This strikes the right balance, lines 67 to 68: "....although greatly impaired.....".

D. Line 149, "Cone signaling to a postsynaptic partner is intact in G369i KI mice". This description is inaccurate. Here there is only WT and KI, and the text reads as follows in line 162: "terminals (Figure 6b). The ON and OFF components of EPSCs in G369i KI HCs were measurable, although lower in amplitude than in WT (Figure 6a,b)." Neither "measurable" nor "lower in amplitude" meet the definition of "intact", and actual numerical values are lacking in the text.

We have added results showing that there are no light responses in the Cav1.4 KO horizontal cells and have modified the sentence to: “Cone synaptic responses are present in horizontal cells of G369i KI but not Cav1.4 KO mice”.

We have modified discussion of these results as (line 210-213): “Consistent with the lack of mature ribbons and abnormal cone pedicles (Fig.1), HC light responses were negligible in Cav1.4 KO mice (Fig.8a,b). In contrast, the ON and OFF responses were present in G369i KI HCs although significantly lower in amplitude than in WT HCs (Fig. 8a,b).”

E. Please add a legend to Figure 6a to indicate the intensities. The shape of the KI responses is different from the control which is worthy of discussion: i) there is no clear cessation of HC EPSCs in the KI during the light ON period (when release stops, Im fluctuations should be minimal), and ii) the "peaked" appearances of the initial 500ms of the On and Off periods are very similar in shape for the KI (hard to interpret in the same fashion as a control response). How were the On and Off amplitudes analyzed? Furthermore, the OFF current is not summarized in Figure 6D, but should not this be when Cav3 should be opening and triggering release: Off response-EPSC? Lastly, Figure 6b,d shows a ~70% reduction in On-current in the KI, and the KI example of 6b an 80% reduction in Off current compared to WT. Yet, the only place asterisks are used to indicate sig diff is the DNQX data within each genotype in Fig 6d. These data cannot be described as showing "intact" KI responses, and the absence of numerical and statistical values needs to be addressed.

New Fig.8a depicting the horizontal cell light responses has been modified to include the legend indicating light intensities. The ON and OFF amplitudes were analyzed as the peak current amplitudes. This information has been added to the legend.

The reviewer is correct in that the OFF response represents the EPSC whereas the ON response represents the decrease in the EPSC with light. To avoid confusion, we changed the y axis label for the averaged data to read ON or OFF “response” rather than “current” in new Fig.8b.

As the reviewer suggests, the more transient nature of the KI response during the light ON period could result from aberrant continuation of vesicular release during the light-induced hyperpolarization of cones in the KI mice, in contrast to the prolonged suppression of release by light which is evident in the WT responses. We speculated on this difference as follows (lines 237-241):

“In addition to its smaller amplitude, the transient nature of the ON response in G369i KI HCs suggested inadequate cessation of cone glutamate release by light (Fig.8b). Slow deactivation of Cav3 channels and/or their activation at negative voltages20 could give rise to Ca2+ signals that support release following light-induced hyperpolarization of G369i KI cones.”

We added astericks to new Fig.8b,d indicating statistical differences and description of the tests in the legend.

F. line 168 the section titled "Light responses of bipolar cells and visual behavior is spared in G369i KI but not Cav1.4 KO mice".

Changed to: “Light responses of bipolar cells and visual behavior are present in G369i KI but not Cav1.4 KO mice”

Last sentence of erg results, 189-190: "These results suggest that cone-to-CBC signaling is intact in G369i KI mice.". "Spared and intact" are not accurate descriptions. The ERG data presented here shows massive differences between WT and the KI, except in the instance of awaves.

This sentence was removed.

As for Figure 6, the results text related to Figure 7a-d does not present real numbers for ERG responses, and there is no indication of significant differences there or in the Figure panels. For instance, in Figure 7b, b-waves are KI are comparable to KO, except at the two highest-intensity flashes that show KI responses ~20% the amplitude of WT. Presentation of KI and KO data on a 6- to 10-fold expanded scale higher than WT can be misleading: a quick read of these Figure panels might make one incorrectly conclude that the KI is intact while the KO is impaired when compared to WT. The Methods section needs more details on the ERG analysis (e.g. any filtering out of oscillatory potentials when measuring b-wave, and what was the allowable range of time-to-peak for b-wave amplitude, etc..).

The vertical scaling of the ERG results in new Fig.10c,d has been changed so as to reflect clearly diminished responses of the KO and KI vs the WT. Further details regarding the ERG analysis was added to the Methods section.

G. Can you point to other studies that have used the "visible platform swim test" used in Figure 7e, f, and specify further how mice were dark/light adapted prior to the recordings?

As referenced in the Methods, original line 674, the methods we used for the swim test were described in our previous study (PMID 29875267). Other studies that have used this assay include PMIDs: 28262416, 26402607.

(2) The Maddox et al 2020 study does not safely address whether rods have a residual T-type Ca2+ current in the Cav 1.4 KO or KI. The study showed that membrane currents measured from rods in the KI and KO retina were distinct from WT, supporting their claim that L-type Ca2+ current is absent in the KI and KO. However, the recordings had shortcomings that challenge the analysis of Ca2+ currents: i) collected at room temp (22-24{degree sign}C), ii) at an unknown distance from the terminal (uncertain voltage clamp), iii) with a very slow voltage ramp rate that is not suitable for probing T-type currents (Figure 1d Maddox 2020, 140 mV over 1 sec: 7msec/1mV), and iv) at a signal-to-noise that does not allow to resolve a membrane current under 1 pA (avg wt rod Ca2+ current was -3.5 pA, and line noise ~1pA peak-to-peak in Maddox 2020). Suggestion: say T-type currents were not probed in Maddox et al 2020, but Davison et al 2022 did not find PCR signal for Cav3.2 in rods.

We disagree that recordings in the Maddox 2020 study were not sufficient to uncover a T-type current. The voltage ramps in that study were not much slower than that of the Davison et al. 2022 study (they used 0.19 mV/ms). Moreover, in new Supp. Fig.S1, we show that like the slower voltage ramp (0.15 mV/ms) used in the prior study of G369i KI rods, the voltage ramps we used in the present study (0.5 mV/ms), which clearly evoke currents with T-type properties in G369i KI cones (Fig.2a,b, Fig.3a,b) do not evoke currents in WT or G369i KI rods.

Minor comments.

(1) Suggestion: add an overview panel to Figure 1 that shows the rod terminals in the KI. The problem is that cropping out the ribbon and active zone signals from rods, to highlight cones, can give the impression that the cones are partially spared in the KI, and the rods are not spared at all. (yet you nicely clarify this in Figure 4 and in the legend and text, etc.).

We chose to modify the legend with this information as in Fig.4 rather than modify the figure.

(2) Mouse wt cone Ca2+ currents look like L-type currents, as do your monkey and squirrel cone recordings, and also much like those of mouse rods (see Figure S5, Hagiwara et al., 2018 or Grabner and Moser 2021). Your pharm data from mice and squirrels further supports your conclusion, and certainly took much effort. Davison et al 2022 J Neurosci showed PCR results that support their claim that a Cav3 current exists in wt cones. Questions: 1) have you tried PCR? 2) Can you offer more details on what Cav3 KO you tried and what antibodies failed to confirm the KO? As the authors know, one complication is that the deletion of one Cav can be compensated for by the expression of a new Cav. There are 3 types of Cav3s and removal of one type may be compensated for by another Cav3.

We have included drop-seq data (new Supp.Fig.S3) implicating Cav3.2 as the main Cav3 subtype in cones and have modified our discussion of these results accordingly. These experiments did not reveal any changes in Cav3 subtype expression in G369i KI vs WT cones.

(3) Lines 95/96- onward, spend more time telling the story. When working out the biophysical and pharmacological behavior of the Ca2+ currents, you might want to initially refer to the membrane current as a membrane current, and then state how your voltage protocols, intra- and extra-cell solutions, and drugs helped you verify 1) L-type and 2) T-type Ca2+ currents.

We have modified the text with more detail.

(4) If data is in hand, add a ramp I-V to Figure S2, which shows the response of the ground squirrel cone. The steps in S2a are excellent for making your point that a transient current is missing, and the bipolar is a great control to illustrate ML218 works. However, a comparison of a squirrel cone ramp to a bipolar ramp response could complete the figure.

See Reponse to #5 below.

(5) Consider moving Supplementary Figures S2 and S3 to the main text; these are highly relevant to the story, novel, and well-executed.

Fig.S2 and S3 were added as new Figs.4,5. The new Fig.4 includes voltage ramps in ground squirrel cones (panel a) to compare with the bipolar data (panel f).

(6) The nice electron microscopy reconstructions are not elaborated on in any detail, and there is no mention of ribbon size. Is the resolution sufficient to estimate ribbon size, the number of synaptic vesicles around the ribbon and in the adjacent cytosol? The images indicate major changes in the morphology of the terminals. Is the glial envelope similar in WT and KI?

Since ribbons were quantified extensively in the confocal analyses in Fig.6, we felt it unnecessary to add this to the EM analysis which focused mainly on aspects of 3D structure (i.e., arrangement of ribbons, postsynaptic wiring, cone pedicle morphology). We added further discussion of the change in morphology of the G369i KI cone pedicle (lines 200-203): “Compared to WT, ribbons in G369i KI pedicles appeared disorganized and were often parallel rather than perpendicular to the presynaptic membrane (Fig.7a-c). Consistent with our confocal analyses (Fig.1), G369i KI cone pedicles extended telodendria in multiple directions rather than just apically (Fig. 7a).”

While we did not opt to characterize the glial envelope in WT cones, we did add an analysis of synaptic vesicles around ribbons to Table 2.

(7) Discussion line 250: "we found no evidence for a functional contribution of Cav3 in our recordings of cones in WT mice (Figures. 2,3), ground squirrels, or macaque (Supplementary Figures S2 and S3).". I would not use "functional" in this context because when comparing your work to Davison et al 2022, they defined functional as a separate response component driven by Cav3. For instance, they examined the influence of their T-type current on exocytosis (by membrane capacitance) and other features like spiking Ca2+ transients. Suggestion: substitute functional with "detectable", and say "we found no detectable Cav currents". Or if you had Ttype staining, but not T-type Ca2+ currents, then say "no functional current even though there is staining...".

We have modified the text as (lines 336-338): “However, in contrast to recordings of WT mouse cone pedicles in a previous study21, we found no evidence for Cav3-mediated currents in somatic recordings of cones in WT mice (Figs.2,3).”

We propose an alternative interpretation of the results in the Davison et al study concerning the conclusion that Cav3 channels contribute to Ca2+ spikes and exocytosis. That study used 100 µM Ni2+ to block a “T-type” contribution to spike activity in cones. In their Figs.4,5, the spikes are suppressed by 100 µM Ni2+ and 10 µM nifedipine, a Cav1 antagonist, and spared by the T-type selective drug Z944. This is problematic for several reasons. First, as shown by the authors

(their Fig.2A1,A2) and others (PMID: 15541900), 100 µM Ni2+ inhibits Cav1-type currents in photoreceptors. Second, Z944 potentiates Cav1 current in their mouse cones (their Fig.2C1,C2). Thus, both reagents are suboptimal for dissecting the contribution of either Cav subtype to spiking activity. With respect to Cav3 channels and exocytosis, these authors interpreted a reduction in exocytosis upon holding at -39 mV compared to at -69 mV as indicating a loss of a T-type driven component of release. However, Cav1 channel inactivation (PMID: 12473074) could lead to the observed reduction in exocytosis at -30 mV.

(8) Additional literature related to your Intro and Discussion. Regarding CSNB2, related mutations of active zone proteins, and what happens to Ca2+ currents when ribbons are deleted, you might want to consider the following studies that measure Ca2+ currents from rods: conditional KO of RIM1/2 (Grabner et al 2015 JN), KO of ELKS1/2 (Hagiwara et al, 2018 JCB), and KO of Ribeye (Grabner and Moser eLife 2021). In these studies, the Cav currents were absent in rods of the ELKS1/2 DKO, strongly reduced (80%) in the RIM1/2DKO, but altered in more subtle ways (activation-inactivation) without significantly changing steady-state Ca2+ current in the Ribeye KO. This does not seem to support some of the arguments you have made in the Introduction and Discussion regarding ribbon size and Ca2+ currents, yet the suggested literature is related to the topic at hand.

A description of these synaptic proteins as potential mediators of the effect of Cav1.4 on ribbon morphogenesis was added to the Discussion, lines 325-327.

(9) Line 129: "Along with the major constituents of the ribbon, CtBP2, and RIBEYE", for clarity Ribeye has two domains, one that is identical to CtBP2 (B-domain) and the unique Ribeye domain (A-domain) that is only expressed at ribbon synapses. And, Piccolino is also embedded in the ribbon (Brandstaetter lab, Wichmann/Moser labs). In other words, Ribeye and Piccolino are the major constituents of the ribbon.

To avoid confusion, we simply mention Ctbp2 and RIBEYE in the context of the corresponding antibodies that were used to label ribbons.

(10) Abstract: consider to rephrase "Ca2+-independent role of Cav1.4" by "Ca2+-permeationindependent role of Cav1.4" or alike

Sentence changed to: “In CSNB2, we propose that Cav3 channels maintain cone synaptic output provided that the nonconducting role of Cav1.4 in cone synaptogenesis remains intact.”

Reviewer #3 (Recommendations For The Authors):

Cav1.4 voltage-gated calcium channels play an important role in neurotransmission at mammalian photoreceptor synapses. Mutations in the CACNA1f gene lead to congenital stationary night blindness that particularly affects the rod pathway. Mouse Cav1.4 knockout and Cav1.4 knockin models suggest that Cav1.4 is also important for the cone pathway. Deletion of Cav1.4 in the knockout models leads to signaling malfunctions and to abundant morphological re-arrangements of the synapse suggesting that the channel not only has a role in the influx of Ca2+ but also in the morphological organization of the photoreceptor synapse. Of note, also additional Cav-channels have been previously detected in cone synapses by different groups, including L-type Cav1.3 (Wu et al., 2007; pmid; Kersten et al., 2020; pmid), and also T-type Cav3.2 (Davison et al., 2021; pmid 35803735).

In order to study a conductivity-independent role of Cav1.4 in the morphological organization of photoreceptor synapses, the authors generated the knockin (KI) mouse Cav1.4 G369i in a previous study (Maddox et al., eLife 2020; pmid 32940604). The Cav1.4 G369i KI channel no longer works as a Ca2+-conducting channel due to the insertion of a glycine in the pore-forming unit (Madox et al. elife 2020; pmid 32940604). In this previous study (Madox et al. elife 2020; pmid 32940604), the authors analyzed Cav1.4 G369i in rod photoreceptor synapses. In the present study, the authors analyzed cone synapses in this KI mouse.

For this purpose, the authors performed a comprehensive set of experimental methods

including immunohistochemistry with antibodies (also with quantitative analyses), electrophysiological measurements of presynaptic Ca2+ currents from cone photoreceptors in the presence/absence of inhibitors of L-type- and T-type- calcium channels, electron microscopy (FIB-SEM), ERG recordings and visual behavior tests of the Cav G369i KI in comparison to the Cav1.4 knockout and wild-type control mice.

The authors found that the non-conducting Cav channel is properly localized in cone synapses and demonstrated that there are no gross morphological alterations (e.g., sprouting of postsynaptic components that are typically observed in the Cav1.4 knockout). These findings demonstrate that cone synaptogenesis relies on the presence of Cav1.4 protein but not on its Ca2+ conductivity. This result, obtained at cone synapses in the present study, is similar to the previously reported results observed for rod synapses (Maddox et al., eLife 2020, pmid 32940604). No further mechanistic insights or molecular mechanisms were provided that demonstrated how the presence of the Cav channels could orchestrate the building of the cone synapse.

We respectfully disagree regarding the mechanistic advance of our study. As indicated by Reviewer 2, a major advance of our study is in providing a mechanism that can explain the longstanding conundrum that congenital stationary night blindness type 2 mutations that would be expected to severely compromise Cav1.4 function do not produce complete blindness. Our study provides an important contrast to the Maddox et al 2020 study in showing that rods and cones respond differentially to loss of Cav1.4 function, which is also relevant to the visual phenotypes of CSNB2. How the presence of Cav1.4 orchestrates cone synaptogenesis is an important topic that is outside the scope of our present study.

In the present study, the authors also propose a homeostatic switch from L-type to (newly occurring) T-type calcium channels in the Cav1.4 G369i KI mouse as a consequence of the deficient calcium channel conductivity in the Cav1.4 G369i Cav1.4 KI mouse. In cones of the Cav1.4 G369i, the high-voltage activated, L-type Ca2+-entry was abolished, in agreement with their previous paper (Maddox et al., eLife 2020, pmid 32940604). The authors found a lowvoltage activated Ca2+ current instead that they assigned to T-type Ca2+-currents based on pharmacological inhibitor experiments. T-type Ca2+-currents/channels were already previously identified in other studies by independent groups and independent techniques

(electrophysiology, RT-PCR, single-cell sequencing) in cones of wild-type mice (Davison et al.,

2021, pmid 35803735; Macosko et al., 2015, pmid 26000488; Williams et al., 2022, pmid 35650675). In the present manuscript (Figures 3a/b), the authors also observed a low-voltage activated, T-type like current in cones of wild-type mice, that is isradipine-resistant and affected by the T-type inhibitor ML218. This finding appears compatible with a T-type-like current in wildtype cones and is consistent with the published data mentioned above, although the authors interpret this data in a different way in the discussion.

Due to the noise inherent in whole cell voltage clamp measurements and some crossover effects in the pharmacology, we cannot completely exclude the presence of a T-type current in WT mouse cones. However, our results very clearly support a conclusion opposite to that stated by the reviewer. Namely, if WT mouse cones have T-type Ca currents, then they are far smaller than those in the Cav1.4 G369i KI and KO cones. In particular, while we identified message for Cav3.2 in WT mouse cones, we were unable to identify a functional T-type current by either voltage clamp measurements or pharmacology. See below for a detailed rebuttal.

This proposal of a homeostatic switch is not convincingly supported in this reviewer's opinion

(for further details, please see below). Furthermore, no data on possible molecular mechanisms were provided that would support such a proposal of a homeostatic switch of calcium channels. No mechanistic/molecular insights were provided for a proposed homeostatic switch between Ltype to T-type channels that the authors propose to occur between wild-type and Cav1.4 G369i as a consequence of conduction-deficient Cav1.4 G369i channels. Is this e.g. based on posttranslational modifications that switch on T-type channels or regulation at the transcriptional level inducing expression of T-type calcium channel or on other mechanisms? The authors remain descriptive with their central hypotheses. No molecular mechanisms/signaling pathways were provided that would support the idea of such a homeostatic switch.

Homeostatic plasticity refers to the maintenance of neuronal function in response to some perturbation in neuronal activity and can result from changes in the expression of ion channel genes (PMID: 36377048, 32747440, 19778903) or regulatory pathways that modulate ion channels (PMID: 15051886, 32492405). We present multiple lines of evidence showing that Cav3 currents appear in cones upon genetically induced Cav1.4 loss of function and can support cone synaptic responses and visual behavior if cone synapse structure is maintained. Our new transcriptomic studies show no difference between levels of Cav3 channel transcripts in WT and G369i KI cones, suggesting that the appearance of the Cav3 currents in G369i KI cones does not result from an increase in Cav3 gene expression. We are currently investigating our transcriptomic dataset to determine if Cav3 regulatory pathways are upregulated in G369i KI cones and will present this in a follow-up study.

The authors show residual photopic signaling in the non-conducting Cav1.4 G369i KI mouse as judged by the recording of postsynaptic currents, ERG recordings and visual behavior tests though in a reduced manner. The residual cone-based signaling could be based on the nonaffected T-type Ca2+ channel conductivity in cone synapses. Given that the L-type current through Cav1.4 is gone in the Cav1.4 G369i KI as previously shown (Maddox et al., 2020, pmid 32940604), the T-type calcium current will remain. However as discussed above, this does not necessarily support the idea of a homeostatic switch.

A major point which we highlighted with new results is that despite the expression of Cav3 transcripts in WT mouse cones, Cav3 channels do not contribute to the cone Ca2+ current. This is at odds with the Davison et al study (PMID: 35803735, see our response to Reviewer 2, pt 7 for caveats of this study), but our results convincingly show that the Cav3 current appears only when Cav1.4 is genetically inactivated. Pharmacological or electrophysiological methods that should reveal the presence of Cav3 currents do not change the properties of the Ca2+ current in cones of WT mice, ground squirrel, or macaque:

• Figs.2-4: Voltage steps to -40 mV (Fig 2e) that activate a sizeable T-current in G369i KI mouse cones produce a negligible transient at pulse onset in WT mouse cones. Similarly, transient currents that are obvious in G369i KI mouse cones during the final step to -30 mV are absent in WT cones. When we block Cav1.4 with isradipine either in cones of WT mice or ground squirrel, the current that remains does not resemble a Cav3 current but rather a scaled down version of the L-type current. ML218, which readily blocks Cav3 channels in HEK293T cells and in G369i KI cones, has only minor effects in cones of WT mice and ground squirrel; these effects of ML218 can be attributed to non-specific actions on Cav1.4 (new Supp.Fig.S2). New Fig.4 (moved from the supplementary data to the main article) clearly shows that the ML218-sensitive current in ground squirrel cones exhibits properties of Cav1.4 not Cav3 channels.

• Figs.2,5: Holding voltages that inactivate Cav3 channels have no effect on the Ca2+ current in cones of WT mice or macaque (recordings of macaque cones were moved from the supplement to the main article as new Fig.5).

In Figure 4 the authors measured an increase in the size of the active zone (as judged by the size of the bassoon cluster) and of the synaptic ribbons in the Cav1.4 G369i. A mechanistic explanation for this phenomenon was not provided and the underlying molecular mechanisms were not unraveled.

The FIB-SEM data uncover some ultrastructural alteration/misalignments of the synaptic ribbons and misalignments of the regular arrangement of the postsynaptic dendrites in the G369i KI mice. Also concerning this observation, the study remains descriptive and does not reveal the underlying mechanisms as it would be expected for eLife.

We respectfully disagree on the descriptive nature of our study and the need for a full characterization of the molecular mechanism underlying the cone synaptic defects in the G369i KI mouse.

An important study in the field (Zanetti et al., Sci. Rep. 2021; pmid 33526839) should be also cited that used a gain-of-function mutation of Cav1.4 to analyze its functional and structural role in the cone pathway.

We have added citation of this paper to the Discussion (lines 354-356).

In conclusion, the study has been expertly performed but remains descriptive without deciphering the underlying molecular mechanisms of the observed phenomena, including the proposed homeostatic switch of synaptic calcium channels. Furthermore, a relevant part of the data in the present paper (presence of T-type calcium channels in cone photoreceptors) has already been identified/presented by previous studies of different groups (Macosko et al., 2015; pmid 26000488; Davison et al., 2021; pmid 35803735; Williams et al., 2022; pmid 35650675). The degree of novelty of the present paper thus appears limited. I think that the study might be better suited in a more specialized journal than eLife.

We thank the reviewer for acknowledging the rigor of our study but disagree with their evaluation regarding the novelty of our work as outlined in our responses above.

Version 2

Abstract

Abstract

In congenital stationary night blindness type 2 (CSNB2)—a disorder involving the Cav1.4 (L-type) Ca2+ channel—visual impairment is mild considering that Cav1.4 mediates synaptic release from rod and cone photoreceptors. Here, we addressed this conundrum using a Cav1.4 knockout (KO) mouse and a knock-in (G369i KI) mouse expressing a non-conducting Cav1.4. Surprisingly, Cav3 (T-type) Ca2+ currents were detected in cones of G369i KI mice and Cav1.4 KO mice but not in cones of wild-type mouse, ground squirrel, and macaque retina. Whereas Cav1.4 KO mice are blind, G369i KI mice exhibit normal photopic (i.e., cone-mediated) visual behavior. Cone synapses, which fail to form in Cav1.4 KO mice, are present, albeit enlarged, and with some errors in postsynaptic wiring in G369i KI mice. While Cav1.4 KO mice lack evidence of cone synaptic responses, electrophysiological recordings in G369i KI mice revealed nominal transmission from cones to horizontal cells and bipolar cells. In CSNB2, we propose that Cav3 channels maintain cone synaptic output provided that the nonconducting role of Cav1.4 in cone synaptogenesis remains intact. Our findings reveal an unexpected form of homeostatic plasticity that relies on a non-canonical role of an ion channel.

Introduction

At the first synapse in the visual pathway, graded electrical signals produced in rod and cone photoreceptors gate the release of glutamate onto postsynaptic neurons. Photoreceptor synapses are specialized with a ribbon organelle, which helps prime synaptic vesicles, 1, 2 and postsynaptic neurites of horizontal and bipolar cells that invaginate deep within the terminal 3. A variety of proteins interact with the ribbon and synaptic vesicles near release sites (i.e., active zones) 4. The importance of these proteins for vision is illustrated by the numerous inherited retinal diseases linked to mutations in their encoding genes 5.

One such gene is CACNA1F, which encodes the voltage-gated Ca2+ (Cav) channel expressed in retinal photoreceptors, Cav1.4 6–8. Among the sub-family of Cav1 L-type channels, Cav1.4 exhibits unusually slow inactivation that is well-matched for supporting the tonic, Ca2+-dependent release of glutamate from photoreceptor synaptic terminals in darkness 9, 10. More than 200 mutations in CACNA1F cause vision disorders including congenital stationary night blindness type 2 (CSNB2) 11, 12. These mutations are broadly categorized as producing a gain of function or loss of function in Cav1.4 13. How these mutations in CACNA1F lead to the variable clinical phenotypes of CSNB2 is largely unknown. Symptoms may include strabismus, low visual acuity, and in many cases, night blindness 14, 15. The latter suggests a primary defect in rod pathways, which is surprising given that Cav1.4 knockout (KO) mice are completely blind and lack any evidence of either rod or cone synaptic responses 6, 7, 16. A major caveat is that rod and cone synapses do not form in Cav1.4 KO mice 7, 16–18. Thus, Cav1.4 KO mice are not suitable for studies of how CACNA1F mutations differentially affect rod and cone pathways or for efforts to uncover how the biophysical properties of Cav1.4 shape photoreceptor synaptic release properties.

Here, we overcome this hurdle with a knock-in mouse strain (G369i KI) expressing a non-conducting mutant form of Cav1.4 19. We show that, although greatly impaired, cone synapses and downstream signaling through cone pathways can support photopic visual function in G369i KI mice. This novel mechanism requires the ability of the Cav1.4 protein, independent of its Ca2+ conductance, to nucleate the assembly of cone ribbon synapses and involves an aberrant Cav3 (T-type) conductance that appears when Cav1.4 Ca2+ signals are compromised.

Results

Cone pedicles and ribbons are present in G369i KI mice but not in Cav1.4 KO mice

The G369i mutation results in an insertion of a glycine residue in a transmembrane domain, which prevents Ca2+ permeation through Cav1.419. The presence of rod synapses in G369i KI mice but not Cav1.4 KO mice indicates that the Cav1.4 protein and not its Ca2+ conductance is required for rod synapse assembly19. To test if this is also true for cones, we performed immunofluorescence and confocal analyses using antibodies against cone arrestin (CAR) and CtBP2 to label cone terminals (i.e., pedicles) and ribbons, respectively (Fig. 1). In Cav1.4 KO mouse retinas, cone pedicles were shrunken and retracted into the outer nuclear layer (ONL, Fig. 1a) with little evidence of elongated ribbons (Fig. 1b). In contrast, cone pedicles in G369i KI mice were normally localized in the outer plexiform layer (OPL, Fig. 1a) and were populated by multiple ribbons that were associated with labeling for Cav1.4 (Fig. 1b). Some abnormalities were apparent in G369i KI cone pedicles, such as extremely elongated ribbons and telodendria extending towards the inner nuclear layer (Fig.1b). These results indicate that Cav1.4 Ca2+ signals are dispensable for the integrity of ribbons and the pedicle but are necessary for their overall refinement.

Cone pedicles and the mutant Cav1.4 G369i channel are normally localized in the retina of G369i KI mice but not Cav1.4 KO mice.

Confocal images of the outer nuclear layer (ONL) and outer plexiform layer (OPL) of retina from WT, G369i KI, and Cav1.4 KO mice that were labeled with antibodies against cone arrestin (CAR), CtBP2, and Cav1.4. a, Inverted images of CAR labeling. Lower panels depict pedicles labeled by CAR antibodies (dotted outlines) and correspond to the boxed regions in the upper panels. Cone pedicles remain in the OPL of WT and G369i KI retina (arrows) but are misshapen and retracted into the ONL of the CaV1.4 KO retina (open arrowheads). Solid arrowheads indicate telodendria which extend only apically in WT retina but extend basally and laterally in G369i KI retina. b, Deconvolved confocal images showing CaV1.4 labeling near cone ribbons in WT and G369i KI pedicles (arrows) and ribbon spheres without CaV1.4 labeling in the CaV1.4 KO pedicle (arrowheads). Rod-associated CtBP2 and Cav1.4 labeling was removed for clarity.

Cav3 channel activity is present in cones of G369i KI and Cav1.4 KO mice but not WT mice, ground squirrel, or macaque retina

A prevailing yet unsupported hypothesis regarding the relatively mild visual phenotypes in CSNB2 is that additional Cav subtypes may compensate for Cav1.4 loss-of-function in cones. If so, then Ca2+ currents (ICa) mediated by these subtypes should be evident in cones of Cav1.4 KO and G369i KI mice. To test this, we performed patch clamp recordings of cones in retinal slices of adult WT, G369i KI, and Cav1.4 KO mice under conditions designed to isolate ICa. During voltage ramps in WT cones, ICa activated around −50 mV and peaked near −20 mV, consistent with the properties of Cav1.4 (Fig. 2a-d). While not observed in rods of G369i KI or Cav1.4 KO mice19, a small-amplitude ICa that activated around −60 mV and peaked near −35 mV was detected in cones of these mice (Fig.2a-d). To uncover this low voltage-activated ICa, we used a faster voltage ramp (0.5 mV/ms) than in the previous study of G369i KI rods (∼0.15 mV/ms)19. With this faster voltage ramp, ICa was not apparent in G369i KI rods (Supp.Fig.S1). Therefore, the aberrant ICa is an adaptation to Cav1.4 loss-of-function in cones but not rods.

A low-voltage activated ICa is present in cones of G369i KI and CaV1.4 KO but absent in cones of WT mice.

a, Representative traces of ICa evoked by voltage ramps. b, ICa traces from a normalized to their peak current amplitude to illustrate the hyperpolarizing shift in the I-V from cones of G369i KI and CaV1.4 KO mice. c,d, I-V (c) and G-V (d) relationships for ICa evoked by 50 ms, +5 mV increments from a holding voltage of −90 mV. Vm, test voltage. WT, n = 13; G369i KI, n = 9; CaV1.4 KO, n = 3. e, Representative ICa traces and voltage protocol (bottom) for steady-state inactivation. ICa was evoked by a conditioning pre-pulse from −90 mV to various voltages for 500 ms followed by a test pulse to −30 mV for 50 ms. f, Steady-state inactivation data from cones as recorded in e. I/Imax represents the current amplitude of each −30-mV test pulse (I) normalized to current amplitude of the −30-mV test pulse preceded by a −90 mV conditioning pre-pulse (Imax) and was plotted against pre-pulse voltage. Shaded region indicates window current for G369i KI cones. Line without symbols represents G-V curve for G369i KI cones replotted from d. WT, n = 8; G369i KI, n = 8; Cav1.4 KO, n = 3. In graphs c,d, and f, smooth lines represent Boltzmann fits, symbols and bars represent mean ± SEM, respectively. g, Representative ICa traces from WT and G369i KI cones and voltage protocol (top). ICa was evoked by a step from - 100 mV to −30 mV before (P1) or after (P2) a 500 ms step to −60 mV. h,i, Graphs depicting peak amplitude of ICa during P1 and P2 steps (h) and the ratio of the amplitudes of ICa evoked by P2 and P1 pulses (P2/P1; i) from cones recorded as in g. WT, n = 6; G369i KI, n = 6, *p < 0.05 by paired t-test.

We next used step depolarizations to identify the type of Cav channel underlying ICa in G369i KI and Cav1.4 KO cones. Analysis of current-voltage (I-V) and conductance-voltage (G-V) relationships revealed a significant, hyperpolarizing shift (∼10 mV) in the voltage of half maximal activation (Vh) in G369i KI cones (Fig.2c,d; Table 1). ICa in G369i KI cones was not sustained as in WT cones but inactivated rapidly during 500-ms step depolarizations (Fig. 2e). For G369i KI cones, overlay of the conductance-voltage (Fig. 2d) and inactivation curves revealed a sizeable window current (Fig. 2f). These features of ICa in G369i KI and Cav1.4 KO cones resembled those of Cav3 T-type channels rather than Cav1.420.

In a previous study, a Cav3-like current was detected in patch clamp recordings of cone pedicles in WT mouse cones21. However, in recordings of WT mouse cone somas, we did not observe a low-voltage activated component in the I-V or G-V curves that would be indicative of a Cav3 subtype (Fig.2c,d). To further test for a Cav3 contribution to ICa in WT mouse cones, we used a double pulse voltage protocol where ICa was evoked before (P1) and after (P2) a 500-ms step to −60 mV (Fig.2g). With this protocol, the inactivation of Cav3 channels by the −60 mV step should lead to a reduction in the amplitude of the P2 vs the P1 current. While such a reduction was clearly evident in cones of G369i KI mice (i.e., P2/P1 < 1), there was no significant difference in the P2 and P1 currents in cones of WT mice (Fig.2g-i).

We also used antagonists selective for Cav1 and Cav3 channels (isradipine and ML 218, respectively). As expected, ICa in WT cones was significantly but incompletely suppressed by isradipine (Fig. 3a,b). The Vh of the I-V was similar before (−36.5 ± 0.8 mV) and after (−38.3 ± 1.8 mV) perfusion of isradipine (t(3,4)=0.90, p = 0.43 by paired t-test), suggesting that the residual ICa after exposure to isradipine arose from unblocked Cav1.4 rather than Cav3 channels. While ML 218 modestly inhibited ICa in some WT cones, the effect was not statistically significant (Fig. 3c,d) and could be attributed to weak activity on Cav1.4. In transfected HEK293T cells, ML 218 modestly inhibited ICa and caused a negative shift in the Vh for Cav1.4, in contrast to its strong inhibition of Cav3.2 (Supp.Fig. S2). Like its effects on Cav3.2, ML 218 nearly abolished ICa in cones of G369i KI mice whereas isradipine had little effect (Fig. 3a-d).

Pharmacological characterization of ICa in cones of WT and G369i KI mice.

a, Representative traces for ICa evoked by voltage ramps in cones of retinas from WT (top) and G369i KI (bottom) mice before (baseline) and after exposure to isradipine (ISR, 1 μM). Middle panel, WT ICa from voltage ramps in the top panel were normalized to their peak ICa to clarify the similar properties of ICa before and after ISR exposure. b, Peak ICa before (−) and during (+) ISR exposure from cones as recorded in a. WT, n = 4; G369i KI, n = 5; *p < 0.05 by paired t-test. c, Representative traces for ICa evoked by voltage ramps in cones of retinas from WT or G369i KI mice before (baseline) and after exposure to ML 218 (5 μM). Middle panel, WT ICa from voltage ramps in the top panel were normalized to their peak ICa to clarify the similar properties of ICa before and after ML 218 exposure. d,

Peak ICa before (−) and during (+) ML 218 exposure. WT, n = 5; G369i KI, n = 5; *p < 0.05 by paired t-test.

Finally, we recorded from ground squirrel and macaque cones, where the large amplitude of ICa facilitates pharmacological and biophysical analyses. Consistent with its actions on Cav1.4 in WT mouse cones (Fig. 3b) and transfected HEK293T cells (Supp.Fig. S2), ML 218 caused an insignificant inhibition of peak ICa (−7.0 ± 20.8%, n = 6 cones) as well as a negative shift in the voltage-dependence of activation in ground squirrel cones (ΔV1/2 = −1.16 ± 0.94 mV, n = 6, mean ± SD, p = 0.029, t-test; Fig. 4a,b). By contrast, application of isradipine dramatically suppressed ICa in cones in various regions of the ground squirrel retina (Fig. 4c,d). Subsequent application of ML 218 did not reveal a Cav3-like current: the I-V and G-V of the isradipine + ML 218-sensitive current were shifted in the positive rather than the negative direction relative to the isradipine-sensitive current (Fig.4c-g). This result is consistent with the time- and voltage-dependent block of Cav1 channels by dihydropyridine antagonists such as isradipine 22. As a positive control, we confirmed that ML 218 blocked a prominent Cav3-type current in ground squirrel type 3a OFF cone bipolar cells (Fig.4h,i). In recordings of macaque cone pedicles, we compared the properties of ICa using a holding voltage of −90 mV or −50 mV, the latter of which should inactivate Cav3 channels (Fig.5a). There was no difference in the I-V or G-V relationships obtained at these holding voltages (Fig.5b-e). As in WT mouse cones (Fig.2g-i), there was also no difference in ICa recorded before and after an inactivating pulse to −60 mV (Fig.5f,g). We conclude that Cav3 channels do not normally contribute to ICa in cones of WT mice, ground squirrel, and macaque.

Pharmacological characterization of ICa in ground squirrel cones.

a, Representative traces corresponding to baseline-corrected ICa evoked by voltage ramps before (baseline) and during application of ML 218. b, G-V relationship of ICa in a. c, Representative traces corresponding to ICa evoked by voltage ramps in control (baseline) and in ISR (2 μM) alone or in ISR+ML 218 (5 μM). d, Peak ICa from cones including the record from c before (−) and during (+) ISR or ISR+ML 218 block. The changes during the addition of ML218 were small but significant (sup., n = 7, ISR: p < 0.0001, ML218: p = 0.0047; mid., n = 5, ISR: p < 0.0001, ML218: p = 0.0001; inf., n = 5, ISR: p = 0.0019, ML218: p = 0.0460; two-tailed t test) and were likely due to a further non-specific, time-dependent reduction in the Cav1 current. e, ICa was evoked by steps from −85 mV to voltages between −65 and −5 mV in increments of +10 mV (protocol shown below the current traces in f). Representative traces are shown in control, ISR, or ISR+ML 218. Current traces during steps to −55 and −45 mV lacked both transient and ML 218-senstive components. Dashed lines indicate zero current. f, Traces show the ICa sensitive to ISR (top) and ISR+ML218 (middle). The ICa recorded in ISR was subtracted from the baseline ICa (top). ICa recorded in ISR+ML218 was subtracted from ICa recorded in ISR (middle). g, Top, I-V relationship of peak and steady-state ICa from f. Bottom, G-V relationship of the ICa blocked by ISR and ISR + ML218. Smooth lines represent Boltzmann fits. Symbols and bars represent mean ± SEM (n = 4 cones). Due a negative shift in the activation properties caused by ML 218 on a residual Cav1 current that is assumed to remain at the end of the experiment, the G-V curve of the ICa isolated through subtraction by applying ISR+ML 281 underwent a statistically insignificant shift to the right. Data presented in e-g are from the same cone. h, Representative ICa traces from voltage ramps in an OFF cb3a bipolar cell before and during the application of ML 218. V1/2 was shifted to the right by 13.1 ± 4.3 mV (n = 3 cb3a cells, mean ± SD) consistent with the block of a Cav3-like conductance. In cb3b OFF bipolar cells, ML 218 produced only a slight leftward shift in an ICa that had a more depolarized activation range, consistent with the exclusive expression a Cav1-type current (ΔV1/2 = −1.9 ± 0.4 mV, n = 3 cells; data not shown). i, Neurobiotin fill shows cb3a axon stratification and morphology in retina labeled with antibodies against bassoon and choline acetyltransferase (CHAT) to label the OFF and ON sublamina of the IPL.

Characterization of ICa in patch clamp recordings of cone pedicles of macaque retina.

a-c, Representative current traces (a) and corresponding I-V (b) and G-V (c) relationships for ICa evoked by 200-ms steps from a holding voltage of −90 mV or −50 mV. n = 3 cones. d,e, V1/2 (d) and slope factor (e) obtained from Boltzmann fit of data in c. p = 0.50 in d; p = 0.25 in e by Wilcoxon matched pairs signed rank test. f, Voltage protocol (top) and representative ICa traces (bottom left) from cones as recorded in Fig. 2g-i. Bottom right, expanded view of the boxed region in the left traces. g, Peak ICa during P1 and P2 steps from cones recorded as in f. n = 5 cones, p = 0.62 by Wilcoxon matched pairs signed rank test.

To test whether the appearance of the Cav3-mediated ICa in G369i KI cones could result from an increase in the expression of a specific Cav3 subtype (i.e., Cav3.1, Cav3.2, or Cav3.3), we initially tried labeling with commercially available antibodies against Cav3.2. However, the data were unreliable since these antibodies produced a similar labeling pattern in Cav3.2 KO retinal tissue (data not shown). As an alternative, we performed single cell RNA sequencing on cells isolated from the retina of WT and G369i KI mice (Supp. Fig.S3). We sequenced 28,782 cells and, with unsupervised clustering at resolution=0.5, we grouped the cells into 20 clusters. Clusters 8 and 16 were identified as cones based on the expression of cone marker genes (e.g., Opn1sw, Opn1mw, Arr3, Gnat2). Consistent with previous studies21, 23, Cacna1h encoding the Cav3.2 subtype was the major Cav3 transcript expressed in WT mouse cones. There was no significant difference in Cacna1h expression between WT and G369i KI cones according to this analysis (Supp. Fig.S3). Thus, the enhanced activity of Cav3 channels in cones of G369i KI mice results from a mechanism other than increased transcription of Cacna1h.

Cone synapses are enlarged with some errors in postsynaptic wiring in G369i KI mice

The replacement of Cav1.4 with Cav3 as the primary conduit for Ca2+ influx in G369i KI cones allowed us to dissect the importance of Cav1.4 Ca2+ signals for the molecular and structural organization of the cone synapse. To this end, we first analyzed WT and G369i KI cone synapses by immunofluorescence and confocal microscopy. Presynaptic proteins such as bassoon and members of the postsynaptic signaling complex in depolarizing (ON) cone bipolar cells (GPR179, mGluR6, and TRPM124) were enriched near cone ribbons in G369i KI mice as in WT mice (Fig. 6a). Compared to WT mice, the labeled structures in G369i KI mice occupied a larger volume (Fig. 6b,c) but were of lower intensity (Fig. 6d), suggesting an increase in the spread rather than in the levels of synaptic proteins. Unlike in WT mice, the volume of presynaptic and postsynaptic proteins increased linearly with the volume of the pedicle in G369i KI mice (Fig. 6e-i). These results show that the molecular determinants of the cone synapse can assemble in G369i KI mice but change in their sub-synaptic distribution in ways that correlate with the expansion of the cone pedicle.

Immunofluorescence characterization of cone synapses in WT and G369i KI mice.

a, Confocal images of the OPL of WT and G369i KI mice labeled with antibodies against cone arrestin (CAR) and proteins that are presynaptic (CtBP2, bassoon) or postsynaptic (GPR179, TRPM1, mGluR6). Every other panel shows high-magnification, deconvolved images of single pedicles labeled with cone arrestin (rod spherule-associated signals were removed for clarity). Arrows indicate ribbon synapses, which appear enlarged in the G369i KI pedicles. b, Violin plot representing volume of presynaptic protein labeling normalized to the volume of their respective CAR-labeled pedicles (volume occupancy). c, Violin plot representing volume of postsynaptic protein labeling normalized to the volume of their respective CAR-labeled pedicles. d, Violin plot representing fluorescence intensity of postsynaptic proteins. For b-d, **p < 0.01, ***p < 0.001, ****p < 0.0001, unpaired t-tests. e-i, Dependence of synapse size on pedicle size. Volumes corresponding to labeling of CtBP2 (e: p = 0.051, r = 0.4 for WT; p < 0.0001, r = 0.88 for G369i KI), CaV1.4 (f: p = 0.8, r = 0.06 for WT; p = 0.002, r = 0.88 for G369i KI), bassoon (g: p = 0.1, r = 0.34 for WT; p < 0.0001, r = 0.75 for G369i KI), mGluR6 (h: p = 0.32, r = 0.35 for WT; p = 0.002, r = 0.73 for G369i KI) and TRPM1 (i: p = 0.32, r = 0.35 for WT; p = 0.007, r = 0.66 for G369i KI) are plotted against pedicle volume. Dashed and solid lines represent fits by linear regression for WT and G369i KI, respectively.

The cone synapse is structurally complex, with the dendritic tips of two horizontal cells and an intervening ON cone bipolar cell invaginating deeply into the pedicle near the ribbon 3. To test how the switch in Cav subtypes might affect this arrangement of postsynaptic partners, we generated 3D reconstructions of cone synapses by serial block-face scanning electron microscopy (SBFSEM; Fig. 7). As with the enlarged synaptic contacts (Fig. 6), structural modifications in G369i KI pedicles were evident. Compared to WT, ribbons in G369i KI pedicles appeared disorganized and were often parallel rather than perpendicular to the presynaptic membrane (Fig.7a-c). Consistent with our confocal analyses (Fig.1), G369i KI cone pedicles extended telodendria in multiple directions rather than just apically (Fig. 7a). In addition, a slightly larger fraction of synaptic sites in the G369i KI pedicle (35% vs 14% in WT) formed incorrect postsynaptic partnerships including with glia (Fig.7b-d’; Table 2). Nevertheless, the number of ribbons is normal in G369i KI cones and about half of the ribbons make invaginating contacts with the appropriate cell types (i.e., both horizontal cells and cone bipolar cells; Table 2). On average, there were more vesicles associated with ribbons in cones of G369i KI mice (245 per ribbon) than of WT mice (178 per ribbon; Table 2), but the difference did not reach statistical significance. Together with our immunofluorescence data, these results indicate that cone synapses undergo relatively subtle morphological changes in G369i KI mice, perhaps in response to the loss of Cav1.4 Ca2+ signals.

Three-dimensional reconstruction of cone synapses by SBFSEM.

Reconstructions were made of WT and G369i KI pedicles (n = 2 each). a, 3D renderings showing ribbons (magenta) within cone pedicles (gray) from WT and G369i KI mice. b, 3D renderings show ribbon sites in a WT cone pedicle contacting one horizontal (HC1; yellow) and three bipolar cells (BC1-3; purple). b’, Raw image from b shows a single plane example of BC1-2 and HC1 contacting the ribbon site. c-d’, Single plane raw (c,d) images and 3D reconstructions (c’,d’) show ribbon sites within the G369i KI cone pedicle contacting in c,c’: horizontal cells (HC1-2) only (upper panels), CBCs (BC1-2) only (lower panels); and in d,d’: a glial cell (orange) and an unknown partner. The glial cell completely envelops the pedicle. Inset in d’ shows glial-contacting ribbon site (arrow). In other panels, arrows indicate points of contact between ribbons and other postsynaptic elements.

Postsynaptic responses to light are present in horizontal cells of G369i KI mice but not Cav1.4 KO mice

Compared to Cav1.4 KO mice, the relative preservation of cone synapses in G369i KI mice allowed the unique opportunity to test whether a Cav subtype other than Cav1.4 could support synaptic release. To this end, we analyzed synaptic transmission between cones and horizontal cells (HCs). In darkness, glutamate released from cones depolarizes horizontal cells via activation of α-amino-3-hydroxy-5-methyl-4-isoxazolepropionic (AMPA)/kainate receptors25, 26. The resulting excitatory postsynaptic current (EPSC) undergoes a decline in response to light stimuli that hyperpolarize cones27. In WT HCs, a 1-s light flash (λ=410 nm) inhibited the standing EPSC, which is reflected as an outward (hyperpolarizing “ON”) current (Fig. 8a). Upon termination of the light, an inward (depolarizing “OFF”) current signaled the resumption of the EPSC (Fig. 8a). Both the ON and OFF responses in WT HCs increased with light intensity, reflecting the impact of luminance on the presynaptic membrane potential of cones and the subsequent change in glutamate release from their terminals (Fig. 8b).

Patch clamp recordings of light responses of horizontal cells in WT, G369i KI and Cav1.4 KO mice.

a-b, Representative traces from whole-cell patch clamp recordings of horizontal cells held at −70 mV (a) and quantified data (b) for currents evoked by 1 s light flashes (λ = 410 nm) plotted against light intensities. In b, peak current amplitudes during (ON Response) and after (OFF Response) the light stimuli were plotted against photon flux per μm2 (Φq/μm2). Data represent mean ± SEM. WT, n = 8; G369i KI, n = 13; Cav1.4 KO, n = 8. There was a significant difference in responses of WT, G369i KI, and Cav1.4 KO horizontal cells. For ON responses, F(16, 216) = 22, p <0.0001 by two-way repeated measures ANOVA, Tukey’s multiple comparisons post-hoc test. For OFF responses, F(16, 216) = 13.61, p < 0.0001 by two-way repeated measures ANOVA, Tukey’s multiple comparisons post-hoc test. c-d, Representative traces from horizontal cells held at −70 mV (c) and quantified data (d) for currents evoked by 1-s light flashes (λ= 410 nm, 1.2 x 105 Φq/μm2) before, during, and after washout of DNQX (20 μM). In d, symbols represent responses from individual cells, n = 5 cells for WT and 3 cells for G369i KI, bars represent mean ± SEM. **, p < 0.01 by paired t-tests.

Consistent with the lack of mature ribbons and abnormal cone pedicles (Fig.1), HC light responses were negligible in Cav1.4 KO mice (Fig.8a,b). In contrast, the ON and OFF responses were present in G369i KI HCs although significantly lower in amplitude than in WT HCs (Fig. 8a,b). As expected, the EPSC in darkness and the light responses were abolished by the AMPA/kainate receptor antagonist, DNQX, in both WT and G369i KI HCs (Fig. 8c,d). In addition to its smaller amplitude, the transient nature of the ON response in G369i KI HCs suggested inadequate cessation of cone glutamate release by light (Fig.8b). Slow deactivation of Cav3 channels and/or their activation at negative voltages20 could give rise to Ca2+ signals that support release following light-induced hyperpolarization of G369i KI cones.

To test if the diminished HC light responses correlated with lower presynaptic Ca2+ signals in G369i KI cones, we performed 2-photon imaging of cone pedicles filled with the Ca2+ indicator Fluo3 and labeled with Alexa-568-conjugated peanut agglutinin. In WT cones, depolarization by K+ (50 mM) caused a sustained increase of Ca2+ (ΔF/F0 = 1.85 ± 0.1, n = 11 terminals) that was significantly lower after the addition of isradipine (ΔF/F0 = 1.44 ± 0.07, p = 0.0015 by repeated measures one-way ANOVA with Tukey’s multiple comparison posthoc test; F[1.608, 16.08] = 55.17; p < 0.0001; Fig.9a,b). The Ca2+ signal remaining in the presence of isradipine could arise from unblocked Cav1.4 channels and/or Ca2+ release from intracellular stores28. In G369i KI cones, the peak amplitude of the K+-evoked Ca2+ signal (ΔF/F0 = 1.07 ± 0.02, n = 14 terminals) was significantly lower than in WT cones (t(23)=8.74, p < 0.0001 by unpaired t-test, n = 14 terminals). The decrease in the Ca2+ signal in the G369i KI cone pedicles during isradipine exposure paralleled that during the depolarization (Fig.9a), suggesting it was not a consequence of Cav1 blockade. The lower amplitude and more transient Ca2+ signals mediated by Cav3 channels may be cleared more rapidly in G369i KI cones as compared to prolonged Cav1.4-initiated Ca2+ signals in WT cones. Taken together, our results suggest that Cav3 channels nominally support Ca2+ signals and synaptic transmission in cones of G369i KI mice.

Light responses of bipolar cells and visual behavior are present in G369i KI but not Cav1.4 KO mice

While HCs play important roles in lateral inhibition of photoreceptor output, the vertical dissemination of visual information from cones to the inner retina is relayed by glutamate to ON and OFF cone bipolar cells (CBCs) which express mGluR6 and AMPA/kainate receptors, respectively. To test for the existence of cone to CBC transmission in G369i KI mice, we recorded electroretinograms (ERGs) under light-adapted conditions using Cav1.4 KO mice as a negative control (Fig. 10a-d). In these recordings, the light-induced response of photoreceptors and the postsynaptic response of ON CBCs corresponds to the a- and b-waves, respectively. Unlike in Cav1.4 KO mice, the a-waves of G369i KI mice resembled those in WT mice, which indicates that cones do not degenerate in this mouse strain (Fig. 10a,b). While greatly reduced in amplitude, the b-wave was measurable in G369i KI mice and significantly larger than in Cav1.4 KO mice at the highest light intensities (F(14,104)=41, p<0.0001, post-hoc Tukey’s p=0.0038 at 2.3 log cd·s/m2; Fig. 10a,b). We also recorded flicker ERGs using 10 Hz light stimuli that can isolate cone pathways involving both ON and OFF CBCs29, 30. In WT mice, flicker responses exhibited two peaks, one at a lower irradiance (−2 log cd·s/m2) and one at higher irradiance (0.5 log cd·s/m2; Fig.10c,d). The peak at the lower irradiance is attributed to responses in both rod and cone pathways, and the peak at higher irradiance is attributed to responses exclusively in cone pathways29, 30. Flicker responses in G369i KI mice showed a similar non-monotonic relation as in WT mice but were significantly lower in amplitude (F(26,143)=51.18, p<0.0001, post-hoc Tukey’s p=0.0237 at 2.5 log cd·s/m2; Fig.10c,d). The inverted flicker responses at higher illuminations in G369i KI mice were absent in Cav1.4 KO mice and may result from the hyperpolarizing contribution of cone-to-OFF CBC transmission (Fig.10c).

To test whether the cone synaptic responses in G369i KI mice enable vision-guided behavior, we used a swim test that assesses the ability of mice to identify a visible platform31. Under scotopic (dark) conditions, G369i KI and Cav1.4 KO mice took significantly longer to find the platform compared to WT mice (Fig. 10e,f). These results are consistent with flicker response assays (Fig.10c,d) as well as the absence of ICa in rods and rod-to-rod bipolar cell synaptic transmission in G369i KI mice19. Under photopic (light) conditions, G369i KI mice but not Cav1.4 KO mice performed as well as WT mice (Fig. 10e,f). Thus, G369i KI mice have enough visual function under photopic conditions to quickly find the platform. Collectively, our results suggest that Cav3 channels can support cone synaptic responses that are sufficient for visual behavior in G369i KI but not Cav1.4 KO mice.

Discussion

The nervous system has remarkable abilities to adapt to pathological perturbations in neuronal activity. In the retina, ablation or degeneration of photoreceptors triggers various postsynaptic mechanisms that maintain some level of visual function in rod or cone pathways. These include remodeling of bipolar cell dendrites and their synapses32–34 as well as changes in the sensitivity of bipolar cells to photoreceptor input and inhibitory modulation34, 35. To our knowledge, this study provides the first evidence for a presynaptic form of homeostatic plasticity that originates within photoreceptors. Using Cav1.4 KO and G369i KI mice, we identify the upregulation of a Cav3 conductance as a common response to Cav1.4 loss-of-function in cones. However, Cav3 channels can only compensate for Cav1.4 loss-of-function if the basic aspects of cone synapse structure are maintained through expression of the G369i mutant channel. Thus, our results also highlight a crucial, non-conducting role for the Cav1.4 protein that allows cone synapses to function in the absence of Cav1.4 Ca2+ signals.

A non-canonical role for Cav1.4 in regulating cone synapse assembly

As shown for rod synapses19, ribbons and other components of the pre- and post-synaptic complex assemble at cone synapses in G369i KI mice (Figs. 1,6,7). Cav3 Ca2+ signals are dispensable for this process since their presence in Cav1.4 KO cones is not accompanied by any semblance of ribbon synapses (Fig.1b). An intimate relationship between Cav1.4 and ribbons is supported by the colocalization of Cav1.4 and puncta resembling ribbon precursor spheres in the developing OPL7. Cav1.4 interacts directly or indirectly with a variety of ribbon synapse associated proteins including RIM 36, CAST/ERC2 37, bassoon38, unc11939–41, and a major component of the ribbon, RIBEYE42. Cav1.4 could pioneer sites of ribbon assembly perhaps by serving as a docking or nucleation site for the active zone. At the same time, ribbon-associated proteins help cluster Cav1.4 near ribbons since knockout of these proteins in mice leads to fewer presynaptic Cav1.4 channels and in some cases shortened ribbons43. Regardless of the mechanism, our findings show that the formation of photoreceptor synaptic complexes does not require Ca2+ influx through Cav1.4 channels.

Functional Cav3 channels are present in cones in the absence of Cav1.4 Ca2+ signals

Our drop-seq analysis aligns with previous studies indicating that Cacna1h is expressed at extremely low levels compared to Cacna1f in WT mouse cones (Supp.Fig.S3)23, 44. However, in contrast to recordings of WT mouse cone pedicles in a previous study21, we found no evidence for Cav3-mediated currents in somatic recordings of cones in WT mice (Figs.2,3). We do not feel the discrepancy results from the different recording configurations given the electrically compact nature of cones. Moreover, recordings from macaque cone pedicles and from the ellipsoids of cylindrically-shaped ground squirrel cones also did not reveal a Cav3-mediated current (Figs.4,5). We acknowledge that there are caveats of the pharmacological approach used in our study as well as the previous work21. For example, dihydropyridine antagonists such as isradipine have relatively low affinity for some Cav1 subtypes, including Cav1.422, 45 and have strongly voltage-dependent blocking properties. At a holding voltage of −90 mV, isradipine at 1 μM causes only ∼80% inhibition of Cav1.4 with greater block at depolarized voltages22, which agrees with our recordings of cones in WT mice (Fig.3a) and ground squirrels (Fig.4b-e). Moreover, in the low micromolar range, Cav3 blockers such as Z944 46 and ML 218 (Supp.Fig. S2) have nonspecific effects on Cav1 current amplitude and activation voltage. Thus, Cav channels mediating a ICa that is suppressed by Cav3 blockers and spared by dihydropyridines at negative voltages may be mistakenly categorized as Cav3 subtypes. In our experiments, the biophysical properties of the isradipine-sensitive ICa in cones of WT mice and ground squirrels resembled only those of Cav1, whereas those of the ML 218-sensitive ICa in cones of G369i KI mice resembled only those of Cav3. Therefore, we favor the interpretation that Cav3 contributes to ICa in mouse cones only upon silencing of Cav1.4 either by KO of the corresponding protein or by silencing its conductance.

Since Cacna1h expression was similar in WT and G369i KI cones (Supp.Fig.S3), what causes the increased Cav3 channel activity in the latter? Among the Cav3 subtypes, Cav3.2 is uniquely sensitive to post-translational modifications that facilitate the opening of the channel and/or its cell surface trafficking47. For example, Cav3.2 interacts with the deubiquitinating enzyme USP5 which enhances levels of Cav3.2 protein levels by suppressing its proteasomal degradation48. In addition, Cav3.2 is phosphorylated by calmodulin-dependent kinase II49 and cyclin-dependent kinase 550, which enhances channel activation and current density, respectively. Further interrogation of our drop-seq dataset could illuminate novel Cav3.2-regulatory pathways that are modified in G369i KI cones.

Consequences of Cav1.4 loss of function and relevance for CSNB2

Despite having the same complement of proteins as in WT mice, cone synapses in G369i KI mice were enlarged and made errors in postsynaptic partner selection (Figs.6,7). While I-V curves indicate similar peak ICa amplitudes in cones of WT and G369i KI mice near the membrane potential of cones in darkness (−45 to −50 mV51, Fig.2a,c), the strong inactivation of Cav3 channels and/or reduced Ca2+ release from intracellular stores likely account for the greatly diminished Ca2+ signals in G369i KI pedicles (Fig.9). Paradoxically, reductions in presynaptic Ca2+ in rod photoreceptors correlated with illumination-dependent shrinkage of ribbons in vivo and in vitro2, 52, 53. However, ribbon synapse size increases with low presynaptic Ca2+ in zebrafish sensory hair cells54. The mechanism involves decreased mitochondrial Ca2+ uptake and increase in the redox state of nicotinamide adenine dinucleotide (NAD+/NADH ratio) 55. Cav3 Ca2+ signals may decay too quickly in G369i KI cones (Fig.9) to enable mitochondrial Ca2+ uptake or other mechanisms that trim ribbons.

Ca2+ imaging of cone terminals in WT and G369i KI mice.

Depolarization-evoked fluorescence changes (ΔF) were measured by two-photon imaging of Fluo3-filled cone pedicles labeled with PNA-Alexa 568.

a, Representative images at the indicated timepoints (T) before (control), during, and after (wash) bath application of KCl (50 mM) and KCl+ISR (10 μM). For clarity, PNA-Alexa 568 labeling was outlined (magenta, solid circles) and the actual labeling only shown in the initial panel of images. Dashed, yellow circles represent ROIs for analyses. b, Changes in cone pedicle Fluo3 fluorescence were normalized to the initial fluorescence intensity at time= 0 (ΔF/F0). Solid lines and shaded areas represent the mean ± SEM, respectively. WT, n = 11 pedicles; G369i KI, n = 14 pedicles.

An intriguing result was the linear increase in ribbon synapse size with pedicle volume in G369i KI but not in WT mice (Fig. 6e-i), which could represent a form of homeostatic synaptic scaling as has been demonstrated at the neuromuscular junction56, 57. Similarly, the presence of some non-invaginating contacts with incorrect partner pairings in G369i KI mice (Fig.7, Table 2) could be an attempt to increase synaptic connectivity to compensate for weakened synaptic output. Consistent with this possibility, diminished presynaptic Ca2+ signals and non-invaginating cone synapses also characterize the retina of mice lacking the extracellular α2δ−4 subunit of Cav1.458. The cone synapses of G369i KI mice and α2δ−4 KO mice are clearly functional given that photopic visual behavior is preserved in both mouse strains (Fig.10e,f)58 and, in the case of G369i KI mice, HCs exhibit EPSCs whose amplitude decreases with light and AMPA/kainate receptor blockade (Fig.8). Defects in cone synaptic transmission that correlate with aberrant cone synapse morphology and connectivity also typify a mouse strain with a gain-of function mutation in Cav1.4 linked to CSNB259, 60. Thus, balanced presynaptic Ca2+ signals may be needed to encode the correct post-synaptic wiring and structure of the cone synapse.

In addition to structural abnormalities in cone synapses, the inability of Cav3 channels to support sustained presynaptic Ca2+ entry likely contributed to the impaired light responses in HCs and CBCs in G369i KI mice (Figs.8,10). Due to their slow activation and strong inactivation 61, Cav3 channels are not expected to fuel the Ca2+ nanodomains that support fast and sustained components of release, both of which occur only at ribbon sites in cones1 62. Cav3 channels may also be located further from the ribbon than Cav1.4, thus lowering the efficiency of coupling to exocytosis. Regardless of the features that differentiate the contributions of Cav3 and Cav1.4 to cone synaptic release, our findings provide further support for the widely held view that Cav1 channels are well-suited to support transmission at ribbon-type synapses.

Based on extremely heterogeneous clinical presentations, CSNB2 manifests as a spectrum of visual disorders that originate from various mutations in CACNA1F 14, 63. Even though some CSNB2 mutant Cav1.4 channels may traffic normally to the plasma membrane, many of these mutations are expected to produce non-functional, non-conducting Cav1.4 channels64, 65. Yet, the visual phenotypes of CSNB2 patients are not as severe as the complete blindness in Cav1.4 KO mice, which lack any Cav1.4 protein expression and exhibit no signs of visual behavior (Fig.10e,f)58. Collectively, our results suggest that G369i KI mice accurately model Cav1.4 channelopathies in CSNB2 patients that are characterized by a greater impairment in rod than in cone pathways. This interpretation is supported by our findings that G369i KI mice exhibit horizontal cell responses to bright but not dim illumination (Fig. 8), ERG responses under conditions of light adaptation (Fig.10a-d) but not dark-adaptation 19, and visual behavior under photopic but not scotopic conditions (Fig. 10e,f). Together with the enlargement of synaptic sites, modest levels of synaptic release from cones of G369i KI mice may support nominal transmission of visual information through cone pathways. G369i KI mice could also exhibit homeostatic alterations in the inner retina, which are known to support visual function when photoreceptor input is severely compromised 33–35. Future studies of the retinal circuitry and visual behavior of G369i KI mice could identify compensatory pathways that are recruited upon Cav1.4 loss-of-function and how they might be targeted in novel therapies for CSNB2 and related disorders.

Characterization of ERGs and visual behavior in WT, G369i KI, and Cav1.4 KO mice.

a, Representative traces of photopic ERGs recorded in the presence of background green light (20 cd · s/m2) in WT, G369i KI and CaV1.4 KO mice. Flash intensities are shown at left. Arrows and arrowheads depict the a- and b-waves, respectively. b, a-wave (left) and b-wave (right) amplitudes are plotted against light intensity. Symbols and bars represent mean ± SEM. WT, n = 7; G369i KI, n = 5; CaV1.4 KO, n = 6. *, p < 0.05; **, p < 0.01; ***, p < 0.001; ****, p < 0.0001; ns, not significant; Two-way ANOVA with Tukey’s posthoc multiple comparisons. c,d, Representative traces (c) and quantified data (d) for 10 Hz flicker responses evoked by white light flashes of increasing luminance (from −4 to 2 log cd· s/m2). Arrows in c depict inverted waveform responses in G369i KI mice that are absent in Cav1.4 KO mice. Symbols and bars represent mean ± SEM. WT, n = 7; G369i KI, n = 5; CaV1.4 KO, n = 6. *, p < 0.05; **, p < 0.01. Two-way ANOVA with Tukey’s posthoc multiple comparisons. e, Representative swim path traces of WT, G369i KI and CaV1.4 KO mice from the visible platform swim tests performed in the dark (scotopic, upper panel) and light (photopic, lower panel). f, Time required to reach the platform (latency to platform) was compared for each mouse strain. Symbols represent the average of the last 3 swim trials for each mouse of each genotype for both dark and light conditions. Dotted lines represent the mean. WT, n = 11; G369i KI, n = 10; CaV1.4 KO, n = 9. *, p < 0.05; ***, p < 0.001; ****, p < 0.0001; Kruskal-Wallis one-way ANOVA with Dunn’s post hoc analysis.

eLife Assessment

Based on analyses of retinae from genetically modified mice, and from wild-type ground squirrel and macaque, employing microscopic imaging, electrophysiology, and pharmacological manipulations, this valuable study on the role of Cav1.4 calcium channels in cone photoreceptor cells (i) shows that the expression of a Cav1.4 variant lacking calcium conductivity supports the development of cone synapses beyond what is observed in the complete absence of Cav1.4, and (ii) indicates that the cone pathway can partially operate even without calcium flux through Cav1.4 channels, thus preserving behavioral responses under bright light. The evidence for the function of Cav1.4 protein in synapse development is convincing and in agreement with a closely related earlier study by the same authors on rod photoreceptors. The mechanism of compensation of Cav1.4 loss by Cav3 remains unclear but appears to involve post-transcriptional processes. As congenital Cav1.4 dysfunction can cause stationary night blindness, this work relates to a wide range of neuroscience topics, from synapse biology to neuro-ophthalmology.

Joint Public Review

Cav1.4 calcium channels control voltage-dependent calcium influx at photoreceptor synapses, and congenital loss of Cav1.4 function causes stationary night blindness CSNB2. Based on a broad portfolio of methodological approaches - genetic mouse models, immunolabeling and microscopic imaging, serial block-face-SEM, ERGs, and electrophysiology - the authors show that cone photoreceptor synapse development is strongly perturbed in the absence of Cav1.4 protein, and that expression of a nonconducting Cav1.4 channel mitigates these perturbations. Further data indicate that Cav3 channels are present, which, according to the authors, may compensate for the loss of Cav1.4 calcium currents and thus maintain cone synaptic transmission. These data, which are in agreement with a similar study by the same authors on rod photoreceptor synapses, help to explain what functional defects exactly cause CSNB2 and why it is accompanied by only mild visual impairment.

The strengths of the present study are its conceptual and experimental soundness, the broad spectrum of cutting-edge methodological approaches pursued, and the convincing differential analysis of mutant phenotypes. Weaknesses mainly concern the mechanism by which Cav3 channels might partially compensate for the loss of Cav1.4 calcium currents.

Author response:

The following is the authors’ response to the previous reviews.

Intro.

47-48 rewrite sentence

This sentence has been rewritten as: Photoreceptor synapses are specialized with a vesicle-associated ribbon organelle and postsynaptic neurites of horizontal and bipolar cells that invaginate deep within the terminal

Results

Major comment. Lines 100-103

The new rod data presented here looks like an n = 1. Neither the Results section nor Supp Fig S1, describe the number of cells used. Nor do the authors offer a statistical description with averages, etc.. In addition, the single traces are much improved over their previous study (Maddox et al eLife 2020), but the authors have not described any new approach or trick that improved their rod Ica. Neither Methods section nor Supp section describes the procedure for patching rods (solutions, or Vh which is critical for assessing T-type currents).

Suggestion, if more data exists, then present it. Otherwise, drop the argument.

The recording methodology for recording rods was like that for cones and this has been clarified in the Methods section (lines 725-752). Averaged data (n= at least 5 per group) and statistical analyses have been added to Fig.S1 (renamed Figure 2-Figure Supplement 1), and clearly show that no Ca2+ currents are present in the KI rods.

Supp Fig S2. The legend needs to be fixed. Conversion to PDF file may have created these formatting errors.

This has been corrected (renamed Figure 3-Figure Supplement 2).

Fig 8 a. The position of the light stimulus bar in the KO panel appears to be out of place, shifted too far to the left.

This has been corrected.

Major comments. 219-221

The use of Fluo3-AM is not properly stated here. The text reads "cone pedicles filled with the Ca2+ indicator Fluo3". The wording used could be wrongly interpreted as: whole-cell filling of the cones via patch electrode. However, the Methods section describes bathing the retina in Fluo3-AM, which presumably fills PRs, HCs tips, Mueller glia and bpc dendrites. The Results section should acknowledge that the retina was loaded with Fluo3-AM.

The cell types, and their processes (Muellers, HCs, bpc, PRs), present in a cone pedicle ROI will likely contribute to the Fluo3 readout of Ca2+ in the OPL, because 1) the EM images in Fig 7 highlight how interdigitated the processes are with the presynapse, 2) all express Cav channels, and many if not all express L-Type Cavs in their processes (glia, HC, on-bcs and PRs), and 3) all are depolarized with the addition of high extracellular KCl. The inclusion of Isradipine will inhibit L-type Cavs on pre- and post-synaptic targets, failing to specifically isolate PR Ca2+. Furthermore, Glu Receptor blockers are used here, which would be a great idea if the cones were stimulated with light; however, KCl bypasses the excitatory synaptic pathway and depolarizes all processes within the ROI. Hence, all cellular parts in the ROI will potentially contribute to Fluo3-Ca2+ signals.

Suggestions for presentation of these findings. Ultimately your conclusion is suitable " 233 to 234...... Taken together, our results suggest that Cav3 channels nominally support Ca2+ signals and synaptic transmission in cones of G369i KI mice". The dramatic reduction in Fluo3-Ca2+ signals in the OPL G369i retinas (Fig 9) is a valuable finding for the following reasons: 1) the results do not show a clear compensation from intracellular stores that could potentially supersede the T-type currents in the G369i (which is an argument you make), and 2) there is a massive loss of Ca2+ influx in the OPL of G369i retinas. Since G369i is specific to the PRs, and only cones are present in the mutant G369i, the loss of Fluo3-Ca2+ signal in the mutant ROI reflects in large part loss of cone Fluo3-Ca2+ signals. Your findings illustrate the severity of the mutation, which has also been addressed in the various electro-physio sections of the MS.

Figure 9 also needs to be more clear about 1) the loading of the cells with AM-dye, and 2) the presence of glia, HCs and bc dendrites in the PNA demarcated ROIs.

We regret that we did not make this more clear, but our Fluo 3 loading protocol of whole retina followed by vertical slicing allowed for loading primarily of photoreceptors in the portion of the outer retina that we imaged. We clarified this with the following edit to the text (lines 220-226):

“To test if the diminished HC light responses correlated with lower presynaptic Ca2+ signals in G369i KI cones, we performed 2-photon imaging of vertical slices prepared from whole retina that was incubated with the Ca2+ indicator Fluo3-AM and Alexa-568-conjugated peanut agglutinin (PNA) to demarcate regions of interest (ROIs) corresponding to cone pedicles. With this approach Fluo3 fluorescence was detected only in photoreceptors and ganglion cells and not inner retinal cell-types (e.g., horizontal cells, bipolar cells, Mueller cell soma). Thus, Ca2+ signals reported by Fluo3 fluorescence near PNA-labeling originated primarily from cones.”

We also note that given the considerably larger volume of the cone pedicle relative to the postsynaptic neurites of horizontal and bipolar cells, as well as neighboring glia, it seems unlikely that the latter would contribute significantly to the isradipine-sensitive Ca2+ signal measured in the ROI above the PNA labeling. Moreover, to our knowledge the contribution of Cav1 L-type channels to postsynaptic Ca2+ signals in the dendritic tips of horizontal cells and bipolar cells has not been demonstrated.

Version 3

Abstract

Abstract

In congenital stationary night blindness type 2 (CSNB2)—a disorder involving the Cav1.4 (L-type) Ca2+ channel—visual impairment is mild considering that Cav1.4 mediates synaptic release from rod and cone photoreceptors. Here, we addressed this conundrum using a Cav1.4 knockout (KO) mouse and a knock-in (G369i KI) mouse expressing a non-conducting Cav1.4. Surprisingly, Cav3 (T-type) Ca2+ currents were detected in cones of G369i KI mice and Cav1.4 KO mice but not in cones of wild-type mouse, ground squirrel, and macaque retina. Whereas Cav1.4 KO mice are blind, G369i KI mice exhibit normal photopic (i.e., cone-mediated) visual behavior. Cone synapses, which fail to form in Cav1.4 KO mice, are present, albeit enlarged, and with some errors in postsynaptic wiring in G369i KI mice. While Cav1.4 KO mice lack evidence of cone synaptic responses, electrophysiological recordings in G369i KI mice revealed nominal transmission from cones to horizontal cells and bipolar cells. In CSNB2, we propose that Cav3 channels maintain cone synaptic output provided that the nonconducting role of Cav1.4 in cone synaptogenesis remains intact. Our findings reveal an unexpected form of homeostatic plasticity that relies on a non-canonical role of an ion channel.

Introduction

At the first synapse in the visual pathway, graded electrical signals produced in rod and cone photoreceptors gate the release of glutamate onto postsynaptic neurons. Photoreceptor synapses are specialized with a vesicle-associated ribbon organelle and postsynaptic neurites of horizontal and bipolar cells that invaginate deep within the terminal 1. A variety of proteins interact with the ribbon and synaptic vesicles near release sites (i.e., active zones) 2. The importance of these proteins for vision is illustrated by the numerous inherited retinal diseases linked to mutations in their encoding genes 3.

One such gene is CACNA1F, which encodes the voltage-gated Ca2+ (Cav) channel expressed in retinal photoreceptors, Cav1.4 4–6. Among the sub-family of Cav1 L-type channels, Cav1.4 exhibits unusually slow inactivation that is well-matched for supporting the tonic, Ca2+-dependent release of glutamate from photoreceptor synaptic terminals in darkness 7, 8. More than 200 mutations in CACNA1F cause vision disorders including congenital stationary night blindness type 2 (CSNB2) 9, 10. These mutations are broadly categorized as producing a gain of function or loss of function in Cav1.4 11. How these mutations in CACNA1F lead to the variable clinical phenotypes of CSNB2 is largely unknown. Symptoms may include strabismus, low visual acuity, and in many cases, night blindness 12, 13. The latter suggests a primary defect in rod pathways, which is surprising given that Cav1.4 knockout (KO) mice are completely blind and lack any evidence of either rod or cone synaptic responses 4, 5, 14. A major caveat is that rod and cone synapses do not form in Cav1.4 KO mice 5, 14–16. Thus, Cav1.4 KO mice are not suitable for studies of how CACNA1F mutations differentially affect rod and cone pathways or for efforts to uncover how the biophysical properties of Cav1.4 shape photoreceptor synaptic release properties.

Here, we overcome this hurdle with a knock-in mouse strain (G369i KI) expressing a non-conducting mutant form of Cav1.4 17. We show that, although greatly impaired, cone synapses and downstream signaling through cone pathways can support photopic visual function in G369i KI mice. This novel mechanism requires the ability of the Cav1.4 protein, independent of its Ca2+ conductance, to nucleate the assembly of cone ribbon synapses and involves an aberrant Cav3 (T-type) conductance that appears when Cav1.4 Ca2+ signals are compromised.

Results

Cone pedicles and ribbons are present in G369i KI mice but not in Cav1.4 KO mice

The G369i mutation results in an insertion of a glycine residue in a transmembrane domain, which prevents Ca2+ permeation through Cav1.417. The presence of rod synapses in G369i KI mice but not Cav1.4 KO mice indicates that the Cav1.4 protein and not its Ca2+ conductance is required for rod synapse assembly17. To test if this is also true for cones, we performed immunofluorescence and confocal analyses using antibodies against cone arrestin (CAR) and CtBP2 to label cone terminals (i.e., pedicles) and ribbons, respectively (Fig. 1). In Cav1.4 KO mouse retinas, cone pedicles were shrunken and retracted into the outer nuclear layer (ONL, Fig. 1a) with little evidence of elongated ribbons (Fig. 1b). In contrast, cone pedicles in G369i KI mice were normally localized in the outer plexiform layer (OPL, Fig. 1a) and were populated by multiple ribbons that were associated with labeling for Cav1.4 (Fig. 1b). Some abnormalities were apparent in G369i KI cone pedicles, such as extremely elongated ribbons and telodendria extending towards the inner nuclear layer (Fig.1b). These results indicate that Cav1.4 Ca2+ signals are dispensable for the integrity of ribbons and the pedicle but are necessary for their overall refinement.

Cone pedicles and the mutant Cav1.4 G369i channel are normally localized in the retina of G369i KI mice but not Cav1.4 KO mice.

Confocal images of the outer nuclear layer (ONL) and outer plexiform layer (OPL) of retina from WT, G369i KI, and Cav1.4 KO mice that were labeled with antibodies against cone arrestin (CAR), CtBP2, and Cav1.4. a, Inverted images of CAR labeling. Lower panels depict pedicles labeled by CAR antibodies (dotted outlines) and correspond to the boxed regions in the upper panels. Cone pedicles remain in the OPL of WT and G369i KI retina (arrows) but are misshapen and retracted into the ONL of the CaV1.4 KO retina (open arrowheads). Solid arrowheads indicate telodendria which extend only apically in WT retina but extend basally and laterally in G369i KI retina. b, Deconvolved confocal images showing CaV1.4 labeling near cone ribbons in WT and G369i KI pedicles (arrows) and ribbon spheres without CaV1.4 labeling in the CaV1.4 KO pedicle (arrowheads). Rod-associated CtBP2 and Cav1.4 labeling was removed for clarity.

Cav3 channel activity is present in cones of G369i KI and Cav1.4 KO mice but not WT mice, ground squirrel, or macaque retina

A prevailing yet unsupported hypothesis regarding the relatively mild visual phenotypes in CSNB2 is that additional Cav subtypes may compensate for Cav1.4 loss-of-function in cones. If so, then Ca2+ currents (ICa) mediated by these subtypes should be evident in cones of Cav1.4 KO and G369i KI mice. To test this, we performed patch clamp recordings of cones in retinal slices of adult WT, G369i KI, and Cav1.4 KO mice under conditions designed to isolate ICa. During voltage ramps in WT cones, ICa activated around −50 mV and peaked near −20 mV, consistent with the properties of Cav1.4 (Fig. 2a-d). While not observed in rods of G369i KI or Cav1.4 KO mice17, a small-amplitude ICa that activated around −60 mV and peaked near −35 mV was detected in cones of these mice (Fig.2a-d). To uncover this low voltage-activated ICa, we used a faster voltage ramp (0.5 mV/ms) than in the previous study of G369i KI rods (∼0.15 mV/ms)17. With this faster voltage ramp, ICa was not apparent in G369i KI rods (Fig.2 Fig.Supp. 1). Therefore, the aberrant ICa is an adaptation to Cav1.4 loss-of-function in cones but not rods.

A low-voltage activated ICa is present in cones of G369i KI and CaV1.4 KO but absent in cones of WT mice.

a, Representative traces of ICaevoked by voltage ramps. b, ICa traces from a normalized to their peak current amplitude to illustrate the hyperpolarizing shift in the I-V from cones of G369i KI and CaV1.4 KO mice. c,d, I-V (c) and G-V (d) relationships for ICaevoked by 50 ms, +5 mV increments from a holding voltage of −90 mV. Vm, test voltage. WT, n = 13; G369i KI, n = 9; CaV1.4 KO, n = 3. e, Representative ICa traces and voltage protocol (bottom) for steady-state inactivation. ICa was evoked by a conditioning pre-pulse from −90 mV to various voltages for 500 ms followed by a test pulse to −30 mV for 50 ms. f, Steady-state inactivation data from cones as recorded in e. I/Imaxrepresents the current amplitude of each - 30-mV test pulse (I) normalized to current amplitude of the −30-mV test pulse preceded by a −90 mV conditioning pre-pulse (Imax) and was plotted against pre-pulse voltage. Shaded region indicates window current for G369i KI cones. Line without symbols represents G-V curve for G369i KI cones replotted from d. WT, n = 8; G369i KI, n = 8; Cav1.4 KO, n = 3. In graphs c,d, and f, smooth lines represent Boltzmann fits, symbols and bars represent mean ± SEM, respectively. g, Representative ICa traces from WT and G369i KI cones and voltage protocol (top). ICa was evoked by a step from −100 mV to −30 mV before (P1) or after (P2) a 500 ms step to −60 mV. h,i, Graphs depicting peak amplitude of ICa during P1 and P2 steps (h) and the ratio of the amplitudes of ICa evoked by P2 and P1 pulses (P2/P1; i) from cones recorded as in g. WT, n = 6; G369i KI, n = 6, *p < 0.05 by paired t-test.

We next used step depolarizations to identify the type of Cav channel underlying ICa in G369i KI and Cav1.4 KO cones. Analysis of current-voltage (I-V) and conductance-voltage (G-V) relationships revealed a significant, hyperpolarizing shift (∼10 mV) in the voltage of half maximal activation (Vh) in G369i KI cones (Fig.2c,d; Table 1). ICa in G369i KI cones was not sustained as in WT cones but inactivated rapidly during 500-ms step depolarizations (Fig. 2e). For G369i KI cones, overlay of the conductance-voltage (Fig. 2d) and inactivation curves revealed a sizeable window current (Fig. 2f). These features of ICain G369i KI and Cav1.4 KO cones resembled those of Cav3 T-type channels rather than Cav1.418.

In a previous study, a Cav3-like current was detected in patch clamp recordings of cone pedicles in WT mouse cones19. However, in recordings of WT mouse cone somas, we did not observe a low-voltage activated component in the I-V or G-V curves that would be indicative of a Cav3 subtype (Fig.2c,d). To further test for a Cav3 contribution to ICa in WT mouse cones, we used a double pulse voltage protocol where ICa was evoked before (P1) and after (P2) a 500-ms step to −60 mV (Fig.2g). With this protocol, the inactivation of Cav3 channels by the −60 mV step should lead to a reduction in the amplitude of the P2 vs the P1 current. While such a reduction was clearly evident in cones of G369i KI mice (i.e., P2/P1 < 1), there was no significant difference in the P2 and P1 currents in cones of WT mice (Fig.2g-i).

We also used antagonists selective for Cav1 and Cav3 channels (isradipine and ML 218, respectively). As expected, ICa in WT cones was significantly but incompletely suppressed by isradipine (Fig. 3a,b). The Vh of the I-V was similar before (−36.5 ± 0.8 mV) and after (−38.3 ± 1.8 mV) perfusion of isradipine (t(3,4)=0.90, p = 0.43 by paired t-test), suggesting that the residual ICa after exposure to isradipine arose from unblocked Cav1.4 rather than Cav3 channels. While ML 218 modestly inhibited ICa in some WT cones, the effect was not statistically significant (Fig. 3c,d) and could be attributed to weak activity on Cav1.4. In transfected HEK293T cells, ML 218 modestly inhibited ICa and caused a negative shift in the Vhfor Cav1.4, in contrast to its strong inhibition of Cav3.2 (Fig.3-Fig.Supp.2). Like its effects on Cav3.2, ML 218 nearly abolished ICain cones of G369i KI mice whereas isradipine had little effect (Fig. 3a-d).

Pharmacological characterization of ICa in cones of WT and G369i KI mice.

a, Representative traces for ICa evoked by voltage ramps in cones of retinas from WT (top) and G369i KI (bottom) mice before (baseline) and after exposure to isradipine (ISR, 1 μM). Middle panel, WT ICa from voltage ramps in the top panel were normalized to their peak ICa to clarify the similar properties of ICa before and after ISR exposure. b, Peak ICa before (−) and during (+) ISR exposure from cones as recorded in a. WT, n = 4; G369i KI, n = 5; *p < 0.05 by paired t-test. c, Representative traces for ICaevoked by voltage ramps in cones of retinas from WT or G369i KI mice before (baseline) and after exposure to ML 218 (5 μM). Middle panel, WT ICa from voltage ramps in the top panel were normalized to their peak ICa to clarify the similar properties of ICa before and after ML 218 exposure. d, Peak ICa before (−) and during (+) ML 218 exposure. WT, n = 5; G369i KI, n = 5; *p < 0.05 by paired t-test.

Finally, we recorded from ground squirrel and macaque cones, where the large amplitude of ICa facilitates pharmacological and biophysical analyses. Consistent with its actions on Cav1.4 in WT mouse cones (Fig. 3b) and transfected HEK293T cells (Fig.3-Fig.Supp.2), ML 218 caused an insignificant inhibition of peak ICa (−7.0 ± 20.8%, n = 6 cones) as well as a negative shift in the voltage-dependence of activation in ground squirrel cones (ΔV1/2 = −1.16 ± 0.94 mV, n = 6, mean ± SD, p = 0.029, t-test; Fig. 4a,b). By contrast, application of isradipine dramatically suppressed ICain cones in various regions of the ground squirrel retina (Fig. 4c,d). Subsequent application of ML 218 did not reveal a Cav3-like current: the I-V and G-V of the isradipine + ML 218-sensitive current were shifted in the positive rather than the negative direction relative to the isradipine-sensitive current (Fig.4c-g). This result is consistent with the time- and voltage-dependent block of Cav1 channels by dihydropyridine antagonists such as isradipine 20. As a positive control, we confirmed that ML 218 blocked a prominent Cav3-type current in ground squirrel type 3a OFF cone bipolar cells (Fig.4h,i). In recordings of macaque cone pedicles, we compared the properties of ICausing a holding voltage of −90 mV or −50 mV, the latter of which should inactivate Cav3 channels (Fig.5a). There was no difference in the I-V or G-V relationships obtained at these holding voltages (Fig.5b-e). As in WT mouse cones (Fig.2g-i), there was also no difference in ICa recorded before and after an inactivating pulse to - 60 mV (Fig.5f,g). We conclude that Cav3 channels do not normally contribute to ICa in cones of WT mice, ground squirrel, and macaque.

Pharmacological characterization of ICa in ground squirrel cones.

a, Representative traces corresponding to baseline-corrected ICa evoked by voltage ramps before (baseline) and during application of ML 218. b, G-V relationship of ICa in a. c, Representative traces corresponding to ICa evoked by voltage ramps in control (baseline) and in ISR (2 μM) alone or in ISR+ML 218 (5 μM). d, Peak ICa from cones including the record from c before (−) and during (+) ISR or ISR+ML 218 block. The changes during the addition of ML218 were small but significant (sup., n = 7, ISR: p < 0.0001, ML218: p = 0.0047; mid., n = 5, ISR: p < 0.0001, ML218: p = 0.0001; inf., n = 5, ISR: p = 0.0019, ML218: p = 0.0460; two-tailed t test) and were likely due to a further non-specific, time-dependent reduction in the Cav1 current. e, ICa was evoked by steps from −85 mV to voltages between −65 and −5 mV in increments of +10 mV (protocol shown below the current traces in f). Representative traces are shown in control, ISR, or ISR+ML 218. Current traces during steps to −55 and −45 mV lacked both transient and ML 218-senstive components. Dashed lines indicate zero current. f, Traces show the ICa sensitive to ISR (top) and ISR+ML218 (middle). The ICa recorded in ISR was subtracted from the baseline ICa (top). ICa recorded in ISR+ML218 was subtracted from ICa recorded in ISR (middle). g, Top, I-V relationship of peak and steady-state ICa from f. Bottom, G-V relationship of the ICa blocked by ISR and ISR + ML218. Smooth lines represent Boltzmann fits. Symbols and bars represent mean ± SEM (n = 4 cones). Due a negative shift in the activation properties caused by ML 218 on a residual Cav1 current that is assumed to remain at the end of the experiment, the G-V curve of the ICa isolated through subtraction by applying ISR+ML 281 underwent a statistically insignificant shift to the right. Data presented in e-g are from the same cone. h, Representative ICa traces from voltage ramps in an OFF cb3a bipolar cell before and during the application of ML 218. V1/2 was shifted to the right by 13.1 ± 4.3 mV (n = 3 cb3a cells, mean ± SD) consistent with the block of a Cav3-like conductance. In cb3b OFF bipolar cells, ML 218 produced only a slight leftward shift in an ICa that had a more depolarized activation range, consistent with the exclusive expression a Cav1-type current (ΔV1/2 = −1.9 ± 0.4 mV, n = 3 cells; data not shown). i, Neurobiotin fill shows cb3a axon stratification and morphology in retina labeled with antibodies against bassoon and choline acetyltransferase (CHAT) to label the OFF and ON sublamina of the IPL.

Characterization of ICa in patch clamp recordings of cone pedicles of macaque retina.

a-c, Representative current traces (a) and corresponding I-V (b) and G-V (c) relationships for ICa evoked by 200-ms steps from a holding voltage of −90 mV or −50 mV. n = 3 cones. d,e, V1/2 (d) and slope factor (e) obtained from Boltzmann fit of data in c. p = 0.50 in d; p = 0.25 in e by Wilcoxon matched pairs signed rank test. f, Voltage protocol (top) and representative ICa traces (bottom left) from cones as recorded in Fig. 2g-i. Bottom right, expanded view of the boxed region in the left traces. g, Peak ICa during P1 and P2 steps from cones recorded as in f. n = 5 cones, p = 0.62 by Wilcoxon matched pairs signed rank test.

To test whether the appearance of the Cav3-mediated ICain G369i KI cones could result from an increase in the expression of a specific Cav3 subtype (i.e., Cav3.1, Cav3.2, or Cav3.3), we initially tried labeling with commercially available antibodies against Cav3.2. However, the data were unreliable since these antibodies produced a similar labeling pattern in Cav3.2 KO retinal tissue (data not shown). As an alternative, we performed single cell RNA sequencing on cells isolated from the retina of WT and G369i KI mice (Fig.2-Fig.Supp.3). We sequenced 28,782 cells and, with unsupervised clustering at resolution=0.5, we grouped the cells into 20 clusters. Clusters 8 and 16 were identified as cones based on the expression of cone marker genes (e.g., Opn1sw, Opn1mw, Arr3, Gnat2). Consistent with previous studies19, 21, Cacna1h encoding the Cav3.2 subtype was the major Cav3 transcript expressed in WT mouse cones. There was no significant difference in Cacna1h expression between WT and G369i KI cones according to this analysis (Fig.2-Fig.Supp.3). Thus, the enhanced activity of Cav3 channels in cones of G369i KI mice results from a mechanism other than increased transcription of Cacna1h.

Cone synapses are enlarged with some errors in postsynaptic wiring in G369i KI mice

The replacement of Cav1.4 with Cav3 as the primary conduit for Ca2+ influx in G369i KI cones allowed us to dissect the importance of Cav1.4 Ca2+ signals for the molecular and structural organization of the cone synapse. To this end, we first analyzed WT and G369i KI cone synapses by immunofluorescence and confocal microscopy. Presynaptic proteins such as bassoon and members of the postsynaptic signaling complex in depolarizing (ON) cone bipolar cells (GPR179, mGluR6, and TRPM122) were enriched near cone ribbons in G369i KI mice as in WT mice (Fig. 6a). Compared to WT mice, the labeled structures in G369i KI mice occupied a larger volume (Fig. 6b,c) but were of lower intensity (Fig. 6d), suggesting an increase in the spread rather than in the levels of synaptic proteins. Unlike in WT mice, the volume of presynaptic and postsynaptic proteins increased linearly with the volume of the pedicle in G369i KI mice (Fig. 6e-i). These results show that the molecular determinants of the cone synapse can assemble in G369i KI mice but change in their sub-synaptic distribution in ways that correlate with the expansion of the cone pedicle.

Immunofluorescence characterization of cone synapses in WT and G369i KI mice.

a, Confocal images of the OPL of WT and G369i KI mice labeled with antibodies against cone arrestin (CAR) and proteins that are presynaptic (CtBP2, bassoon) or postsynaptic (GPR179, TRPM1, mGluR6). Every other panel shows high-magnification, deconvolved images of single pedicles labeled with cone arrestin (rod spherule-associated signals were removed for clarity). Arrows indicate ribbon synapses, which appear enlarged in the G369i KI pedicles. b, Violin plot representing volume of presynaptic protein labeling normalized to the volume of their respective CAR-labeled pedicles (volume occupancy). c, Violin plot representing volume of postsynaptic protein labeling normalized to the volume of their respective CAR-labeled pedicles. d, Violin plot representing fluorescence intensity of postsynaptic proteins. For b-d, **p < 0.01, ***p < 0.001, ****p < 0.0001, unpaired t-tests. e-i, Dependence of synapse size on pedicle size. Volumes corresponding to labeling of CtBP2 (e: p = 0.051, r = 0.4 for WT; p < 0.0001, r = 0.88 for G369i KI), CaV1.4 (f: p = 0.8, r = 0.06 for WT; p = 0.002, r = 0.88 for G369i KI), bassoon (g: p = 0.1, r = 0.34 for WT; p < 0.0001, r = 0.75 for G369i KI), mGluR6 (h: p = 0.32, r = 0.35 for WT; p = 0.002, r = 0.73 for G369i KI) and TRPM1 (i: p = 0.32, r = 0.35 for WT; p = 0.007, r = 0.66 for G369i KI) are plotted against pedicle volume. Dashed and solid lines represent fits by linear regression for WT and G369i KI, respectively.

The cone synapse is structurally complex, with the dendritic tips of two horizontal cells and an intervening ON cone bipolar cell invaginating deeply into the pedicle near the ribbon 1. To test how the switch in Cav subtypes might affect this arrangement of postsynaptic partners, we generated 3D reconstructions of cone synapses by serial block-face scanning electron microscopy (SBFSEM; Fig. 7). As with the enlarged synaptic contacts (Fig. 6), structural modifications in G369i KI pedicles were evident. Compared to WT, ribbons in G369i KI pedicles appeared disorganized and were often parallel rather than perpendicular to the presynaptic membrane (Fig.7a-c). Consistent with our confocal analyses (Fig.1), G369i KI cone pedicles extended telodendria in multiple directions rather than just apically (Fig. 7a). In addition, a slightly larger fraction of synaptic sites in the G369i KI pedicle (35% vs 14% in WT) formed incorrect postsynaptic partnerships including with glia (Fig.7b-d’; Table 2). Nevertheless, the number of ribbons is normal in G369i KI cones and about half of the ribbons make invaginating contacts with the appropriate cell types (i.e., both horizontal cells and cone bipolar cells; Table 2). On average, there were more vesicles associated with ribbons in cones of G369i KI mice (245 per ribbon) than of WT mice (178 per ribbon; Table 2), but the difference did not reach statistical significance. Together with our immunofluorescence data, these results indicate that cone synapses undergo relatively subtle morphological changes in G369i KI mice, perhaps in response to the loss of Cav1.4 Ca2+ signals.

Three-dimensional reconstruction of cone synapses by SBFSEM.

Reconstructions were made of WT and G369i KI pedicles (n = 2 each). a, 3D renderings showing ribbons (magenta) within cone pedicles (gray) from WT and G369i KI mice. b, 3D renderings show ribbon sites in a WT cone pedicle contacting one horizontal (HC1; yellow) and three bipolar cells (BC1-3; purple). b’, Raw image from b shows a single plane example of BC1-2 and HC1 contacting the ribbon site. c-d’, Single plane raw (c,d) images and 3D reconstructions (c’,d’) show ribbon sites within the G369i KI cone pedicle contacting in c,c’: horizontal cells (HC1-2) only (upper panels), CBCs (BC1-2) only (lower panels); and in d,d’: a glial cell (orange) and an unknown partner. The glial cell completely envelops the pedicle. Inset in d’ shows glial-contacting ribbon site (arrow). In other panels, arrows indicate points of contact between ribbons and other postsynaptic elements.

Postsynaptic responses to light are present in horizontal cells of G369i KI mice but not Cav1.4 KO mice

Compared to Cav1.4 KO mice, the relative preservation of cone synapses in G369i KI mice allowed the unique opportunity to test whether a Cav subtype other than Cav1.4 could support synaptic release. To this end, we analyzed synaptic transmission between cones and horizontal cells (HCs). In darkness, glutamate released from cones depolarizes horizontal cells via activation of α-amino-3-hydroxy-5-methyl-4-isoxazolepropionic (AMPA)/kainate receptors23, 24. The resulting excitatory postsynaptic current (EPSC) undergoes a decline in response to light stimuli that hyperpolarize cones25. In WT HCs, a 1-s light flash (λ=410 nm) inhibited the standing EPSC, which is reflected as an outward (hyperpolarizing “ON”) current (Fig. 8a). Upon termination of the light, an inward (depolarizing “OFF”) current signaled the resumption of the EPSC (Fig. 8a). Both the ON and OFF responses in WT HCs increased with light intensity, reflecting the impact of luminance on the presynaptic membrane potential of cones and the subsequent change in glutamate release from their terminals (Fig. 8b).

Patch clamp recordings of light responses of horizontal cells in WT, G369i KI and Cav1.4 KO mice.

a-b, Representative traces from whole-cell patch clamp recordings of horizontal cells held at −70 mV (a) and quantified data (b) for currents evoked by 1 s light flashes (λ = 410 nm) plotted against light intensities. In b, peak current amplitudes during (ON Response) and after (OFF Response) the light stimuli were plotted against photon flux per μm2 (Φq/μm2). Data represent mean ± SEM. WT, n = 8; G369i KI, n = 13; Cav1.4 KO, n = 8. There was a significant difference in responses of WT, G369i KI, and Cav1.4 KO horizontal cells. For ON responses, F(16, 216) = 22, p <0.0001 by two-way repeated measures ANOVA, Tukey’s multiple comparisons post-hoc test. For OFF responses, F(16, 216) = 13.61, p < 0.0001 by two-way repeated measures ANOVA, Tukey’s multiple comparisons post-hoc test. c-d, Representative traces from horizontal cells held at −70 mV (c) and quantified data (d) for currents evoked by 1-s light flashes (λ = 410 nm, 1.2 × 105 Φq/μm2) before, during, and after washout of DNQX (20 μM). In d, symbols represent responses from individual cells, n = 5 cells for WT and 3 cells for G369i KI, bars represent mean ± SEM. **, p < 0.01 by paired t-tests.

Consistent with the lack of mature ribbons and abnormal cone pedicles (Fig.1), HC light responses were negligible in Cav1.4 KO mice (Fig.8a,b). In contrast, the ON and OFF responses were present in G369i KI HCs although significantly lower in amplitude than in WT HCs (Fig. 8a,b). As expected, the EPSC in darkness and the light responses were abolished by the AMPA/kainate receptor antagonist, DNQX, in both WT and G369i KI HCs (Fig. 8c,d). In addition to its smaller amplitude, the transient nature of the ON response in G369i KI HCs suggested inadequate cessation of cone glutamate release by light (Fig.8b). Slow deactivation of Cav3 channels and/or their activation at negative voltages18 could give rise to Ca2+ signals that support release following light-induced hyperpolarization of G369i KI cones.

To test if the diminished HC light responses correlated with lower presynaptic Ca2+ signals in G369i KI cones, we performed 2-photon imaging of vertical slices prepared from whole retina that was incubated with the Ca2+ indicator Fluo3-AM and Alexa-568-conjugated peanut agglutinin (PNA) to demarcate regions of interest (ROIs) corresponding to cone pedicles. With this approach Fluo3 fluorescence was detected only in photoreceptors and ganglion cells and not inner retinal cell-types (e.g., horizontal cells, bipolar cells, Mueller cell soma). Thus, Ca2+ signals reported by Fluo3 fluorescence near PNA-labeling originated primarily from cones. In WT cones, depolarization by K+ (50 mM) caused a sustained increase of Ca2+ (ΔF/F0 = 1.85 ± 0.1, n = 11 terminals) that was significantly lower after the addition of isradipine (ΔF/F0 = 1.44 ± 0.07, p = 0.0015 by repeated measures one-way ANOVA with Tukey’s multiple comparison posthoc test; F[1.608, 16.08] = 55.17; p < 0.0001; Fig.9a,b). The Ca2+ signal remaining in the presence of isradipine could arise from unblocked Cav1.4 channels and/or Ca2+ release from intracellular stores26. In G369i KI cones, the peak amplitude of the K+-evoked Ca2+ signal (ΔF/F0 = 1.07 ± 0.02, n = 14 terminals) was significantly lower than in WT cones (t(23)=8.74, p < 0.0001 by unpaired t-test, n = 14 terminals). The decrease in the Ca2+ signal in the G369i KI cone pedicles during isradipine exposure paralleled that during the depolarization (Fig.9a), suggesting it was not a consequence of Cav1 blockade. The lower amplitude and more transient Ca2+ signals mediated by Cav3 channels may be cleared more rapidly in G369i KI cones as compared to prolonged Cav1.4-initiated Ca2+ signals in WT cones. Taken together, our results suggest that Cav3 channels nominally support Ca2+ signals and synaptic transmission in cones of G369i KI mice.

Ca2+ imaging of cone terminals in WT and G369i KI mice.

Depolarization-evoked fluorescence changes (ΔF) were measured by two-photon imaging of Fluo3-filled cone pedicles labeled with PNA-Alexa 568. a, Representative images at the indicated timepoints (T) before (control), during, and after (wash) bath application of KCl (50 mM) and KCl+ISR (10 μM). For clarity, PNA-Alexa 568 labeling was outlined (magenta, solid circles) and the actual labeling only shown in the initial panel of images. Dashed, yellow circles represent ROIs for analyses. b, Changes in cone pedicle Fluo3 fluorescence were normalized to the initial fluorescence intensity at time= 0 (ΔF/F0). Solid lines and shaded areas represent the mean ± SEM, respectively. WT, n = 11 pedicles; G369i KI, n = 14 pedicles.

Light responses of bipolar cells and visual behavior are present in G369i KI but not Cav1.4 KO mice

While HCs play important roles in lateral inhibition of photoreceptor output, the vertical dissemination of visual information from cones to the inner retina is relayed by glutamate to ON and OFF cone bipolar cells (CBCs) which express mGluR6 and AMPA/kainate receptors, respectively. To test for the existence of cone to CBC transmission in G369i KI mice, we recorded electroretinograms (ERGs) under light-adapted conditions using Cav1.4 KO mice as a negative control (Fig. 10a-d). In these recordings, the light-induced response of photoreceptors and the postsynaptic response of ON CBCs corresponds to the a- and b-waves, respectively. Unlike in Cav1.4 KO mice, the a-waves of G369i KI mice resembled those in WT mice, which indicates that cones do not degenerate in this mouse strain (Fig. 10a,b). While greatly reduced in amplitude, the b-wave was measurable in G369i KI mice and significantly larger than in Cav1.4 KO mice at the highest light intensities (F(14,104)=41, p<0.0001, post-hoc Tukey’s p=0.0038 at 2.3 log cd·s/m2; Fig. 10a,b). We also recorded flicker ERGs using 10 Hz light stimuli that can isolate cone pathways involving both ON and OFF CBCs27, 28. In WT mice, flicker responses exhibited two peaks, one at a lower irradiance (−2 log cd·s/m2) and one at higher irradiance (0.5 log cd·s/m2; Fig.10c,d). The peak at the lower irradiance is attributed to responses in both rod and cone pathways, and the peak at higher irradiance is attributed to responses exclusively in cone pathways27, 28. Flicker responses in G369i KI mice showed a similar non-monotonic relation as in WT mice but were significantly lower in amplitude (F(26,143)=51.18, p<0.0001, post-hoc Tukey’s p=0.0237 at 2.5 log cd·s/m2; Fig.10c,d). The inverted flicker responses at higher illuminations in G369i KI mice were absent in Cav1.4 KO mice and may result from the hyperpolarizing contribution of cone-to-OFF CBC transmission (Fig.10c).

Characterization of ERGs and visual behavior in WT, G369i KI, and Cav1.4 KO mice.

a, Representative traces of photopic ERGs recorded in the presence of background green light (20 cd · s/m2) in WT, G369i KI and CaV1.4 KO mice. Flash intensities are shown at left. Arrows and arrowheads depict the a- and b-waves, respectively. b, a-wave (left) and b-wave (right) amplitudes are plotted against light intensity. Symbols and bars represent mean ± SEM. WT, n = 7; G369i KI, n = 5; CaV1.4 KO, n = 6. *, p < 0.05; **, p < 0.01; ***, p < 0.001; ****, p < 0.0001; ns, not significant; Two-way ANOVA with Tukey’s posthoc multiple comparisons. c,d, Representative traces (c) and quantified data (d) for 10 Hz flicker responses evoked by white light flashes of increasing luminance (from −4 to 2 log cd· s/m2). Arrows in c depict inverted waveform responses in G369i KI mice that are absent in Cav1.4 KO mice. Symbols and bars represent mean ± SEM. WT, n = 7; G369i KI, n = 5; CaV1.4 KO, n = 6. *, p < 0.05; **, p < 0.01. Two-way ANOVA with Tukey’s posthoc multiple comparisons. e, Representative swim path traces of WT, G369i KI and CaV1.4 KO mice from the visible platform swim tests performed in the dark (scotopic, upper panel) and light (photopic, lower panel). f, Time required to reach the platform (latency to platform) was compared for each mouse strain. Symbols represent the average of the last 3 swim trials for each mouse of each genotype for both dark and light conditions. Dotted lines represent the mean. WT, n = 11; G369i KI, n = 10; CaV1.4 KO, n = 9. *, p < 0.05; ***, p < 0.001; ****, p < 0.0001; Kruskal-Wallis one-way ANOVA with Dunn’s post hoc analysis.

To test whether the cone synaptic responses in G369i KI mice enable vision-guided behavior, we used a swim test that assesses the ability of mice to identify a visible platform29. Under scotopic (dark) conditions, G369i KI and Cav1.4 KO mice took significantly longer to find the platform compared to WT mice (Fig. 10e,f). These results are consistent with flicker response assays (Fig.10c,d) as well as the absence of ICa in rods and rod-to-rod bipolar cell synaptic transmission in G369i KI mice17. Under photopic (light) conditions, G369i KI mice but not Cav1.4 KO mice performed as well as WT mice (Fig. 10e,f). Thus, G369i KI mice have enough visual function under photopic conditions to quickly find the platform. Collectively, our results suggest that Cav3 channels can support cone synaptic responses that are sufficient for visual behavior in G369i KI but not Cav1.4 KO mice.

Discussion

The nervous system has remarkable abilities to adapt to pathological perturbations in neuronal activity. In the retina, ablation or degeneration of photoreceptors triggers various postsynaptic mechanisms that maintain some level of visual function in rod or cone pathways. These include remodeling of bipolar cell dendrites and their synapses30–32 as well as changes in the sensitivity of bipolar cells to photoreceptor input and inhibitory modulation32, 33. To our knowledge, this study provides the first evidence for a presynaptic form of homeostatic plasticity that originates within photoreceptors. Using Cav1.4 KO and G369i KI mice, we identify the upregulation of a Cav3 conductance as a common response to Cav1.4 loss-of-function in cones. However, Cav3 channels can only compensate for Cav1.4 loss-of-function if the basic aspects of cone synapse structure are maintained through expression of the G369i mutant channel. Thus, our results also highlight a crucial, non-conducting role for the Cav1.4 protein that allows cone synapses to function in the absence of Cav1.4 Ca2+ signals.

A non-canonical role for Cav1.4 in regulating cone synapse assembly

As shown for rod synapses17, ribbons and other components of the pre- and post-synaptic complex assemble at cone synapses in G369i KI mice (Figs. 1,6,7). Cav3 Ca2+ signals are dispensable for this process since their presence in Cav1.4 KO cones is not accompanied by any semblance of ribbon synapses (Fig.1b). An intimate relationship between Cav1.4 and ribbons is supported by the colocalization of Cav1.4 and puncta resembling ribbon precursor spheres in the developing OPL5. Cav1.4 interacts directly or indirectly with a variety of ribbon synapse associated proteins including RIM 34, CAST/ERC2 35, bassoon36, unc11937–39, and a major component of the ribbon, RIBEYE40. Cav1.4 could pioneer sites of ribbon assembly perhaps by serving as a docking or nucleation site for the active zone. At the same time, ribbon-associated proteins help cluster Cav1.4 near ribbons since knockout of these proteins in mice leads to fewer presynaptic Cav1.4 channels and in some cases shortened ribbons41. Regardless of the mechanism, our findings show that the formation of photoreceptor synaptic complexes does not require Ca2+ influx through Cav1.4 channels.

Functional Cav3 channels are present in cones in the absence of Cav1.4 Ca2+ signals

Our drop-seq analysis aligns with previous studies indicating that Cacna1h is expressed at extremely low levels compared to Cacna1f in WT mouse cones (Supp.Fig.S3)21, 42. However, in contrast to recordings of WT mouse cone pedicles in a previous study19, we found no evidence for Cav3-mediated currents in somatic recordings of cones in WT mice (Figs.2,3). We do not feel the discrepancy results from the different recording configurations given the electrically compact nature of cones. Moreover, recordings from macaque cone pedicles and from the ellipsoids of cylindrically-shaped ground squirrel cones also did not reveal a Cav3-mediated current (Figs.4,5). We acknowledge that there are caveats of the pharmacological approach used in our study as well as the previous work19. For example, dihydropyridine antagonists such as isradipine have relatively low affinity for some Cav1 subtypes, including Cav1.420, 43 and have strongly voltage-dependent blocking properties. At a holding voltage of −90 mV, isradipine at 1 μM causes only ∼80% inhibition of Cav1.4 with greater block at depolarized voltages20, which agrees with our recordings of cones in WT mice (Fig.3a) and ground squirrels (Fig.4b-e). Moreover, in the low micromolar range, Cav3 blockers such as Z944 44 and ML 218 (Supp.Fig. S2) have nonspecific effects on Cav1 current amplitude and activation voltage. Thus, Cav channels mediating a ICa that is suppressed by Cav3 blockers and spared by dihydropyridines at negative voltages may be mistakenly categorized as Cav3 subtypes. In our experiments, the biophysical properties of the isradipine-sensitive ICain cones of WT mice and ground squirrels resembled only those of Cav1, whereas those of the ML 218-sensitive ICa in cones of G369i KI mice resembled only those of Cav3. Therefore, we favor the interpretation that Cav3 contributes to ICa in mouse cones only upon silencing of Cav1.4 either by KO of the corresponding protein or by silencing its conductance.

Since Cacna1h expression was similar in WT and G369i KI cones (Supp.Fig.S3), what causes the increased Cav3 channel activity in the latter? Among the Cav3 subtypes, Cav3.2 is uniquely sensitive to post-translational modifications that facilitate the opening of the channel and/or its cell surface trafficking45. For example, Cav3.2 interacts with the deubiquitinating enzyme USP5 which enhances levels of Cav3.2 protein levels by suppressing its proteasomal degradation46. In addition, Cav3.2 is phosphorylated by calmodulin-dependent kinase II47 and cyclin-dependent kinase 548, which enhances channel activation and current density, respectively. Further interrogation of our drop-seq dataset could illuminate novel Cav3.2-regulatory pathways that are modified in G369i KI cones.

Consequences of Cav1.4 loss of function and relevance for CSNB2

Despite having the same complement of proteins as in WT mice, cone synapses in G369i KI mice were enlarged and made errors in postsynaptic partner selection (Figs.6,7). While I-V curves indicate similar peak ICa amplitudes in cones of WT and G369i KI mice near the membrane potential of cones in darkness (−45 to −50 mV49, Fig.2a,c), the strong inactivation of Cav3 channels and/or reduced Ca2+ release from intracellular stores likely account for the greatly diminished Ca2+ signals in G369i KI pedicles (Fig.9). Paradoxically, reductions in presynaptic Ca2+ in rod photoreceptors correlated with illumination-dependent shrinkage of ribbons in vivo and in vitro50–52. However, ribbon synapse size increases with low presynaptic Ca2+ in zebrafish sensory hair cells53. The mechanism involves decreased mitochondrial Ca2+ uptake and increase in the redox state of nicotinamide adenine dinucleotide (NAD+/NADH ratio)54. Cav3 Ca2+ signals may decay too quickly in G369i KI cones (Fig.9) to enable mitochondrial Ca2+ uptake or other mechanisms that trim ribbons.

An intriguing result was the linear increase in ribbon synapse size with pedicle volume in G369i KI but not in WT mice (Fig. 6e-i), which could represent a form of homeostatic synaptic scaling as has been demonstrated at the neuromuscular junction55, 56. Similarly, the presence of some non-invaginating contacts with incorrect partner pairings in G369i KI mice (Fig.7, Table 2) could be an attempt to increase synaptic connectivity to compensate for weakened synaptic output. Consistent with this possibility, diminished presynaptic Ca2+ signals and non-invaginating cone synapses also characterize the retina of mice lacking the extracellular α2δ−4 subunit of Cav1.457. The cone synapses of G369i KI mice and α2δ−4 KO mice are clearly functional given that photopic visual behavior is preserved in both mouse strains (Fig.10e,f)57 and, in the case of G369i KI mice, HCs exhibit EPSCs whose amplitude decreases with light and AMPA/kainate receptor blockade (Fig.8). Defects in cone synaptic transmission that correlate with aberrant cone synapse morphology and connectivity also typify a mouse strain with a gain-of function mutation in Cav1.4 linked to CSNB258, 59. Thus, balanced presynaptic Ca2+ signals may be needed to encode the correct post-synaptic wiring and structure of the cone synapse.

In addition to structural abnormalities in cone synapses, the inability of Cav3 channels to support sustained presynaptic Ca2+ entry likely contributed to the impaired light responses in HCs and CBCs in G369i KI mice (Figs.8,10). Due to their slow activation and strong inactivation 60, Cav3 channels are not expected to fuel the Ca2+ nanodomains that support fast and sustained components of release, both of which occur only at ribbon sites in cones61 62. Cav3 channels may also be located further from the ribbon than Cav1.4, thus lowering the efficiency of coupling to exocytosis. Regardless of the features that differentiate the contributions of Cav3 and Cav1.4 to cone synaptic release, our findings provide further support for the widely held view that Cav1 channels are well-suited to support transmission at ribbon-type synapses.

Based on extremely heterogeneous clinical presentations, CSNB2 manifests as a spectrum of visual disorders that originate from various mutations in CACNA1F 12, 63. Even though some CSNB2 mutant Cav1.4 channels may traffic normally to the plasma membrane, many of these mutations are expected to produce non-functional, non-conducting Cav1.4 channels64, 65. Yet, the visual phenotypes of CSNB2 patients are not as severe as the complete blindness in Cav1.4 KO mice, which lack any Cav1.4 protein expression and exhibit no signs of visual behavior (Fig.10e,f)57. Collectively, our results suggest that G369i KI mice accurately model Cav1.4 channelopathies in CSNB2 patients that are characterized by a greater impairment in rod than in cone pathways. This interpretation is supported by our findings that G369i KI mice exhibit horizontal cell responses to bright but not dim illumination (Fig. 8), ERG responses under conditions of light adaptation (Fig.10a-d) but not dark-adaptation 17, and visual behavior under photopic but not scotopic conditions (Fig. 10e,f). Together with the enlargement of synaptic sites, modest levels of synaptic release from cones of G369i KI mice may support nominal transmission of visual information through cone pathways. G369i KI mice could also exhibit homeostatic alterations in the inner retina, which are known to support visual function when photoreceptor input is severely compromised 31–33. Future studies of the retinal circuitry and visual behavior of G369i KI mice could identify compensatory pathways that are recruited upon Cav1.4 loss-of-function and how they might be targeted in novel therapies for CSNB2 and related disorders.

Supporting information